### 2024-06-24 summaries/june-summary.tex
initial commit

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[top=0.60in, bottom=0.80in, left=0.60in, right=0.60in]{geometry}
\usepackage{graphicx}
\usepackage{siunitx}
\usepackage{listings}

\newcommand{\Msun}{\mathrm{M_{\odot}}}
\newcommand{\Rsun}{\mathrm{R_{\odot}}}
\newcommand{\MWD}{M_{\mathrm{WD}}}
\newcommand{\Mstar}{M_{\star}}
\newcommand{\alphace}{\alpha_{\mathrm{CE}}}
\newcommand{\alphath}{\alpha_{\mathrm{th}}}
\newcommand{\alpharec}{\alpha_{\mathrm{rec}}}
\newcommand{\Ebind}{E_{\mathrm{bind}}}
\newcommand{\Egrav}{E_{\mathrm{grav}}}
\newcommand{\Eint}{E_{\mathrm{int}}}
\newcommand{\Eth}{E_{\mathrm{th}}}
\newcommand{\Erec}{E_{\mathrm{rec}}}

\newcommand{\au}{\, \mathrm{au}}

\title{Wide Post-Mass Transfer White-Dwarf Binaries}
\author{Runqiu Ye}
\date{June, 2024}

\begin{document}

\maketitle

\section{Common Envelope}

As argued in \cite{yamaguchi_hi,yamaguchi_lo}, unstable mass transfer is expect to be the main channel to form wide WD+MS binaries. In the MESA model, wide post-CE WD+MS binaries can be formed in certain range of initial separation. We hope to explore similar models in COSMIC.

In \cite{yamaguchi_hi}, the lower limit of WD mass $M_{\mathrm{MD, min}}$ in the observational sample is calculated to be $1.244\Msun \sim 1.418\Msun$. The corresponding mass of MS progenitor is expected to be in the range of $6 \sim 9 \Msun$, and the companion mass has median around $1\Msun$ In \cite{yamaguchi_lo}, the WD mass in the Gaia sample ranges from $0.5\Msun$ to $0.8\Msun$, which corresponds to progenitor mass of $1 \Msun$ to $3 \Msun$. The companions mass $\Mstar$ has a median of $0.85 \Msun$. Hence, we consider high mass systems using a $7\Msun + 1\Msun$ binary and low mass systems using a $1.5\Msun + 0.85\Msun$

\subsection{High mass systems ($7\Msun + 1\Msun$)}
In \cite{yamaguchi_hi}, the authors discussed currently observed wide post-mass transfer binary star systems. The MS progenitor of the WD star is expected to be in the range of $6 \sim 9 \Msun$. Therefore, MESA model of a $7\Msun$ star is run, and the evolution up to AGB is followed. After that, the change of separation through common envelope is calculated with
\[
E_{\mathrm{bind}} = \alphace \left(-\frac{G\MWD \Mstar}{2a_f}+\frac{GM_i \Mstar}{2a_i}\right).
\]
The binding energy on LHS is calculated as
\[
E_{\mathrm{bind}} = E_{\mathrm{grav}} + E_{\mathrm{int}} = \int_{M_{\mathrm{core}}}^{M_{\mathrm{tot}}} -\frac{Gm}{r(m)} + U(m) dm
\]
or
\[
E_{\mathrm{bind}} = E_{\mathrm{grav}} + \alphath E_{\mathrm{th}} + \alpharec E_{\mathrm{rec}}.
\]
The results is presented in Figure \ref{yam_hi}, and we want to compare how this energy budget matches with the binding energy in COSMIC. Usually, the binding energy of the common envelope can be included in a structure parameter $\lambda$ such that 
\[
\Ebind = \frac{G M_i M_{\mathrm{env}}}{\lambda R_i}.    
\]
This $\lambda$ is represented in COSMIC as the \verb|lambdaf| flag in \verb|BSEDict|. Hence, we want to calculate the effective $\lambda$ in COSMIC that matches these binding energy models in MESA.

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{yamaguchi-7+1.png}
    \caption{Results of how final separation $a_f$ after common envelope depends on initial separation $a_i$ for different common envelope efficiencies and energy budgets in \cite{yamaguchi_hi}.}
    \label{yam_hi}
\end{figure}

\subsubsection{Methods and results}
To compare model in COSMIC with model in MESA, we run COSMIC binary simulation of a $7\Msun + 1\Msun$ system. We want to see how the final separation $a_f$ depends on the initial separation $a_i$ and the common envelope parameter $\lambda$. Hence, we vary the initial separation in range $2 \sim 8 \au$ and the common envelope flag \verb|lambdaf| in range $0 \sim -100$, which corresponds to $0 \sim 100$ for $\lambda$. Also, we run these parameters with four different common envelope efficiencies — $\alphace = 1$, $\alphace = 0.9$, $\alphace = 0.6$, and $\alphace = 0.3$.

After simulation is finished, we select the timestep that the system finished common envelope for the first time (\verb|evol_type = 8|). At this moment, if the system is a WD+MS system, we record their separation as the final separation $a_f$. Otherwise, we set $a_f = 0$ and consider this initial condition unable to produce a WD+MS system.

We write the resulting final separations into \verb|csv| files and create four heat maps for the four different common envelope efficiencies, as shown in Figure \ref{res_hi}.

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{7+1results.png}
    \caption{COSMIC results of how final separation $a_f$ depends on initial separation $a_i$ and common envelope flag "lambdaf" for different common envelope efficiencies $\alphace$, where the black and red contour corresponds to $0.01 \au$ and $0.15 \au$ as in Figure \ref{yam_hi} respectively. Outside the black contour no desired system forms.}
    \label{res_hi}
\end{figure}

\subsubsection{Discussions}
Taking a closer look as the heatmap in Figure \ref{res_hi}, we notice a few interesting points.

First of all, as $\alphace$ decreases, the final separation decreases in general. This is expected and can be clearly derived from the formula. Qualitatively, more energy from the orbit is needed to eject the envelope for low $\alphace$, so the resulting $a_f$ is smaller.

In addition, it is worth noting that most of these systems experience two mass transfer phases in COSMIC before reaching WD+MS. That is, two common evenlope phases take places, between which there's a period of time. For some cases \verb|kstar_1| changes during this period.

Also notice that the final separation $a_f$ jumps obviously at about $2.9 \au$, and $7.1 \au$, which can be clearly seen from the graph. After investigating the \verb|bpp| array of these simulations, we find out the jump at about $2.9 \au$ and $7.1 \au$ is likely due to the difference of \verb|kstar_1| at the start of RLOF. For $a_i < 2.9 \au$, RLOF starts when \verb|kstar_1 = 3| (First Giant Branch, GB). For $a_i = 2.9 \au$, RLOF starts when \verb|kstar_1 = 4| (Core Helium Burning, CHeB). For $2.9 \au < a_i < 7.1 \au$, RLOF starts when \verb|kstar_1 = 5|, (Early AGB, E-AGB). For $a_i > 7.1 \au$, RLOF starts when \verb|kstar_1 = 6|, (thermal-pulse AGB, TP-AGB).

\begin{figure}
    \centering
    \includegraphics[width = 0.6\linewidth]{jump-zoom.png}
    \caption{Dependence of final separation $a_f$ on initial separation $a_i$ for $\alphace = 0.3$. Simulation results of $\lambda_f = 60$, $70$, and $80$ are shown. The red circles emphasizes the jump of final separation as initial separation increases.}
    \label{jump-zoom}
\end{figure}

Inspecting more closely, we find that the final separation $a_f$ jumps at about $5.2 \sim 5.4 \au$. This is especially obvious with $\alphace = 0.3$. To illustrate, we zoom in the simulation results of $\alphace = 0.3$, and present the change of final separation $a_f$ as initial separation $a_i$ increases in Figure \ref{jump-zoom}. From here, we can clearly see two jumps — one between $5.1\au$ and $5.2\au$, the other at $5.34\au$. To explain this, we again go to the \verb|bpp| array of these systems. We notice that the first jump is probably because the evolution process changes from \textbf{3-7-8-4-3-7-8-4} to \textbf{3-7-8-7-8-4}. This means that the mass transfer changes from dynamically stable to dynamically unstable, leading to more mass loss during a shorter timescale and thus larger final separation. Also, the second jump is probably because \verb|kstar_1| changes from \verb|8| (Naked Helium Star Hertzsprung Gap) to \verb|9| (Naked Helium Star Giant Branch) at the start of the second common envelope. This changes the binding energy of the common evenlope and thus changes the final separation.

Finally, notice that a small triangular gap in the figure for $\alphace = 1$. This is because \verb|kstar_1| becomes a neutron star and no rows in the \verb|bpp| gets selected.

In conclusion, with a large \verb|lambdaf|, it is possible to create wide post-mass transfer WD+MS systems with final separation $a_f > 0.15 \au$, which is the minimal separation of our observed objects.

\subsubsection{Comparison}
After we have obtained the simulation results for different initial conditions, now we can calculate the effective $\lambda$ that matches with results in \cite{yamaguchi_hi}. We use the following method to build the connection between COSMIC model and MESA model. For each fixed initial selaration $a_i$, we loop through the COSMIC results with the same initial separation, and find the \verb|lambdaf| value that results in a final separation closest to MESA model results, documented in the figures of \cite{yamaguchi_hi}. We record this \verb|lambdaf| as the effective $\lambda$ that matches the binding energy formalism at initial separation $a_i$.

For $\Ebind = \Egrav + \Eint$, the fitting results is presented in Figure \ref{fit_hi}, where each line corresponds to different common envelope efficiencies $\alphace$ and differnt energy budgets (either $\Egrav$ only or $\Egrav + \Eint$).

\begin{figure}
    \centering
    \includegraphics[width=1\linewidth]{7+1fit.png}
    \caption{Effective $\lambda$ fitting results of a $7\Msun + 1\Msun$ system for common envelope structure in MESA. \emph{Left:} MESA results in \cite{yamaguchi_hi} of how final separation $a_f$ depends on initial separation $a_i$. \emph{Center:} COSMIC results of how final separation $a_f$ depends on initial separation $a_i$, produced by the effective $\lambda$. \emph{Right:} effective $\lambda$ in COSMIC that corresponds to the MESA results for each initial separation $a_i$. The red line in the left and middle panel represents $0.15 \au$, the minimal separation for observed wide post-CE WD+MS binaries.}
    \label{fit_hi}
\end{figure}

Notice that there is a small bump in the blue line. This is because at $6 \sim 7 \au$, we cannot reach such low final separation in COSMIC while forming a WD+MS binary at the same time. In the left panel which shows the MESA results documented in \cite{yamaguchi_hi}, we notice that even with low common envelope efficiencies, it is possible to create WD+MS binaries with final separation $>0.15 \au$ in MESA as long as internal energy $\Eint$ is included. However, in the right panel we show that this correspond to very large \verb|lambdaf| in COSMIC ($\lambda > \sim 20$).

To furthur compare with the default $\lambda$ value in COSMIC, we include in the right panel an extra line which represents the default $\lambda$ value in COSMIC. This default $\lambda$ in the COSMIC model is calculated following Appendix A of \cite{claeys2014theoretical}. For all of our case, we have $M_{\mathrm{env}} > 1$. The results is shown in Figure \ref{fit_cmp_hi}. Notice that the dafult $\lambda$ value is at order $\sim 1$, far smaller than the effective $\lambda$ needed to recreate results in \cite{yamaguchi_hi}.

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{7+1cmp.png}
    \caption{Same as Figure \ref{fit_hi}, but with default $\lambda$ in COSMIC model included in the right panel.}
    \label{fit_cmp_hi}
\end{figure}

For a different energy budge with $\Ebind = \Egrav + \alphath \Eth + \alpharec \Erec$, the same results is shown in Figure \ref{fit_cmp_eb_hi}

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{7+1ebcmp.png}
    \caption{Same as Figure \ref{fit_cmp_hi}, but with a different energy budget.}
    \label{fit_cmp_eb_hi}
\end{figure}

\subsection{Low mass systems ($1.5\Msun + 0.85\Msun$)}

In \cite{yamaguchi_lo}, the author discussed the formation of WD+MS binaries through common envelope for low mass systems ($1.5\Msun + 0.85\Msun$). In this paper, the companion mass is set to be the median of observed systems $\Mstar = 0.85\Msun$, and the progenitor mass fot the WD is expected to be $1 \sim 3 \Msun$. Hence, $1\Msun$, $1.5\Msun$, and $3\Msun$ models are run in MESA. For the energy budget, thermal energy $\Eth$ is fully included while the recombination energy $\Erec$ is included in part with a parameter $\alpharec$. That is,
\[
    \Ebind = \Egrav + \Eth + \alpharec \Erec.
\]

The results of MEsA models are shown in Figure \ref{yam_lo}

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{yamaguchi-1.5+0.8.png}
    \caption{Results of how final separation $a_f$ after common envelope depends on initial separation $a_i$ for different common envelope efficiencies and energy budgets in \cite{yamaguchi_lo}. Here, for final separation $a_f <0.01\au$, WD+MS binaries are not possible, while $0.15\au$ is the observed minimal separation and $0.7\au$ is the typical separation.}
    \label{yam_lo}
\end{figure}

We take the same approach as in high mass systems to investigate the effective structural parameter $\lambda$ that matches with the common evenlope model in MESA. We ran COSMIC binary evolution model for a system of $1.5\Msun + 0.85\Msun$ for different initial separation $a_i$ and \verb|lambdaf|, to investigate how they affect the final separation $a_f$. The results in shown in Figure \ref{res_lo}.

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{1.5+0.8results.png}
    \caption{COSMIC results of how final separation $a_f$ depends on initial separation $a_i$ and common envelope flag "lambdaf" for different common envelope efficiencies $\alphace$, where the black and red contour corresponds to $0.01 \au$, $0.15 \au$, and $0.7 \au$ as in Figure \ref{yam_lo} respectively. Outside the black contour no desired system forms since WD+MS binaries are not possible for $a_f < 0.01\au$. $0.15\au$ and $0.7\au$ are the observed minimal separation and the typical separation respectively.}
    \label{res_lo}
\end{figure}

We compare this result with results in \cite{yamaguchi_lo} using the same appraoch as high mass systems. In the paper, the author presents separate results for two different method of calculating the binding energy. One of them contains binding energy of the entire envelope and the other only contains binding energy of the outer envelope. In Figure \ref{fit_cmp_lo_whole}, we compare COSMIC with MESA model that includes binding enregy of the entire envelope. In Figure \ref{fit_cmp_lo_pt}, we compare with with MESA model that includes only binding energy of the outer envelope. Notice that we do not have fitting results for $a_i > 4 \au$. This is because we do not get any WD+MS binaries in COSMIC for such initial separation range.

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{1.5+0.8cmp_whole.png}
    \caption{Effective $\lambda$ fitting results of a $1.5\Msun + 0.85\Msun$ system for common envelope structure in MESA. Binding energy of the \textbf{entire envelope} is considered. \emph{Left:} MESA results in \cite{yamaguchi_lo} of how final separation $a_f$ depends on initial separation $a_i$. \emph{Center:} COSMIC results of how final separation $a_f$ depends on initial separation $a_i$, produced by the effective $\lambda$. \emph{Right:} effective $\lambda$ in COSMIC that corresponds to the MESA results for each initial separation $a_i$. The purple line represents default $\lambda$ calculated using \cite{claeys2014theoretical}. The red line and the blue line in the left and middle panel represents $0.15 \au$ and $0.7 \au$ respectively. $0.15 \au$ is the minimal separation for observed wide post-CE WD+MS binaries while $0.7 \au$ is the typical separation for such systems.}
    \label{fit_cmp_lo_whole}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{1.5+0.8cmp_pt.png}
    \caption{Same as Figure \ref{fit_cmp_lo_pt} but only binding enregy of the \textbf{outer envelope} is included.}
    \label{fit_cmp_lo_pt}
\end{figure}

Based on the figure, we find that for low mass systems, the $\lambda$ does not need to be far larger than default to produce WD+MS systems. This is a significant difference between high mass systems and low mass systems.

\section{Stable Mass Transfer}
After exploring the conditions of WD+MS binaries through common envelope, we investigate the possibility of WD+MS binaries formation through stable mass transfer. Similarly, we investigate the condition of forming both high mass systems ($7\Msun + 1\Msun$) and low mass systems ($1.5\Msun + 0.85\Msun$) through stable mass transfer.

\subsection{High mass systems ($7\Msun + 1\Msun$)}

\subsubsection{Methods and results}
First of all, to ensure stable mass transfer, we modify every entry of the \verb|qcrit_array| in \verb|BSEDict| from the default $0.0$ to $20.0$. This will make the actual mass ratio of our system below this critical value for common envelope thorughout the entire process of mass transfer.

Secondly, to investigate dependence on different accretion efficiency, we modify the \verb|acc_lim| flag in \verb|BSEDict|, that is, the fraction $\beta$ of all transfered materials is accreted by the accretor, or $\dot{M_a} = - \beta\dot{M_d}$. For each accretion efficiency, we run COSMIC binary simulation of a $7\Msun + 1 \Msun$ system for initial separation ranging from $2 \au$ to $8 \au$. After the simulatio is finished, we select the first row, if any, that shows a WD+MS binary from \verb|bpp| for each initial separation, and document the final separation. In this way, we can see how final separation depends on the initial separtion, and we can compare them with the observed minimum separation $0.15\au$.

The results is presented in Figure \ref{stable_hi}. The gap in the center means no WD+MS system forms for that initial separation. In fact, a closer inspection of the \verb|bpp| shows that for these systems, two stars merges. The details of the \verb|bpp| will be discussed in a more detailed manner in the next section.

\begin{figure}
    \centering
    \includegraphics[width=0.75\linewidth]{stable_hi_log.png}
    \caption{COSMIC results of WD+MS binaries after stable mass transfer for $7\Msun + 1\Msun$ systems, with accretion efficiency $\beta = 0$, $\beta = 0.5$, and $\beta = 1$ respectively. \emph{Left:} dependence of final separation $a_f$ of the WD+MS binary on initial separation $a_i$. \emph{Right:} The corresponding companion mass in solar units in the WD+MS binary. The gap in both figure represents that no WD+MS system forms for that initial separation.}
    \label{stable_hi}
\end{figure}

Notice that the majority of these systems have final separation $a_f > 0.15 \au$. However, this is \textbf{not the correct approach}, since the companion mass in our observed wide WD+MS binaries has median mass $\sim 1\Msun$. However, in the simulation results shown in Figure \ref{stable_hi}, the final companion mass is about $4 \Msun$ and $6.5\Msun$ for \verb|acc_lim = 0.5| and \verb|acc_lim = 1.0| respectively. Hence, we need to adjust the  initial companion mass $M_{\star, i}$ and the accretion efficiency \verb|acc_lim|, in order to make the final companion mass $M_{\star, f}$ close to $1\Msun$.

\begin{figure}
    \centering
    \includegraphics[width = 0.75\linewidth]{stable7+1ex.png}
    \caption{COSMIC results of WD+MS binaries after stable mass transfer for two systems — $7\Msun + 0.7\Msun$ with accretion efficiency $\beta = 0.1$, and $7\Msun + 0.5\Msun$ with accretion efficiency $\beta = 0.2$. The left and right panel presents the same content as in Figure \ref{stable_hi}.}
    \label{stable_hi_ex}
\end{figure}

Therefore, we lower the initial companion mass $M_{\star, i}$ and adjust the accretion efficiency \verb|acc_lim| and re-run the simulation. The results for $7\Msun + 0.7\Msun$ with \verb|acc_lim=0.1|, and $7\Msun + 0.5\Msun$ with \verb|acc_lim = 0.2| is shown in Figure \ref{stable_hi_ex}.

After adjusting the initial mass of companion and the accretion efficiency, now the final mass of the companion is about $1\Msun$ and we get systems similar to our observed sample. Notice that when initial separation $a_i$ is in the range of about $2.5 \au$ to $4.5 \au$, the final separation increases rapidly. Also, it is only in this region that the final separation goes larger than the observed minimum $0.15 \au$. Hence, in the next subsection, we focus on this sharp increase and try to expain it qualitatively.

\subsubsection{Discussions}
In the previous section, we mentioned how $7\Msun + 1 \Msun$ systems with \verb|acc_lim = 0|, $7 \Msun + 0.7\Msun$ systems with \verb|acc_lim = 0.1|, and $7 \Msun + 0.5 \Msun$ systems with \verb|acc_lim = 0.2| produces WD+MS binaries with final separation $a_f > 0.15 \au$ and companion mass $\sim 1 \Msun$ in certain range of initial separation $a_i$. From the figure, we can clearly see that when initial separation $a_i$ is in the range of $2.5 \au \sim 4.5\au$, the final separation increases sharply as initial separation increases. In this section, we investigate the reason behind this by focusing on $7\Msun + 1\Msun$ binaries with \verb|acc_lim = 0|.

First of all, let's take a look at how the final separation $a_f$ depends on the initial separation $a_i$ throughout the range of $2 \sim 8 \au$. In general, as shown in Figure \ref{stable_hi_ex}, the dependence of $a_f$ on $a_i$ can be roughly divided into four parts — \textbf{pre-spike, spike, merge, and after-merge}, which corresponds to the part before the sharp increase ($a_i < 2.5 \au$), the part where final separation increases sharply ($2.5 \au < a_i < 3.5 \au$), the part where the binary merges and no desired system is formed ($3.5 \au < a_i < 4.5 \au$), and the part after that ($a_i > 4.5 \au$).

A typical \textbf{pre-spike} \verb|bpp| is presented in the left panel of Figure \ref{pre-spike-spike}. Notice that the binary come into contact (\verb|evol_type = 5|), which follows by a common envelope phase. The same is observed in \textbf{spike} binaries, as shown in the right panel of \ref{pre-spike-spike}.

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{pre-spike-spike.png}
    \caption{COSMIC simulation data of $7 \Msun + 1 \Msun$ systems with accretion efficiency $\beta = 0$ and initial separation $2.30 \au$ and $2.90 \au$ respectively. They represent typical "pre-spike" binaries and "spike" binaries. The blue rectangle marks out the common first phase of evolution in both systems — contact and common evnelope mass transfer.}
    \label{pre-spike-spike}
\end{figure}

Two typical \textbf{spike} \verb|bpp| is presented in Figure \ref{spike}. In the left panel, we notice that the stable mass transfer process in the green rectangle widens the orbit a lot. In the right panel, the extra stable mass transfer process in the red rectangle also significantly widens the orbit. However, in the right panel, the orbit widens without the primary's losing a lot of mass, and the corresponding mass transfer in green does not widen the orbit, even shrinking it.

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{sharp-inc-bpp.png}
    \caption{COSMIC simulation of systems $7 \Msun + 1 \Msun$ with accretion efficiency $\beta = 0$ and initial separation $2.80 \au$ and $2.90 \au$ respectively. The red rectangle emphasizes the extra mass transfer process that significantly widens the orbit without the primary losing significant mass.}
    \label{spike}
\end{figure}

To qualitatively explain thses results, note first that for a stable and non-conservative and the mass, we have
\[
\frac{\dot J}{J} \propto (1-\beta) q_d \frac{\dot{M_d}}{M_T}.
\]
While $a \sim J^2$, the non-conservative mass transfer should causes the orbit to expand. Note also that mass transfer begins when the primary is a Naked Helium Star on the Hertzsprung Gap (\verb|evol_type = 8|), whose wind mass loss is significant. Also, for wind mass loss, we have
\[
\frac{\dot a}{a} = - \frac{\dot M}{M_T}.
\] 
Therefore, strong wind stripping mass from the primary will cause the orbit to expand even further. To illustrate this more clearly, we change the \verb|windflag| in \verb|BSEDict| to 0, which correspond to weak wind. Now the $a_i = 2.80 \au$ system has final separation $a_f = 0.05 \au$, and the $a_i = 2.90 \au$ system has final separation $a_f = 0.14 \au$, both smaller than the original $0.09 \au$ and $0.27\au$ that correspond to stronger wind.

From the equations, it is not difficult to see that the increase of separation is more obvious for larger separation $a$ for both wind mass loss and stable, non-conservative mass transfer. Therefore, we conclude qualitatively that for \textbf{pre-spike} and \textbf{spike} systems, non-conservative mass transfer and wind mass loss widens the orbit. While this is more obvious for large separation $a$, this accounts for the spike of final separation $a_f$ as the initial separation $a$ increases.

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{merge-after-merge.png}
    \caption{COSMIC simulation of systems $7 \Msun + 1 \Msun$ with accretion efficiency $\beta = 0$ and initial separation $3.25 \au$ and $5.00 \au$ respectively. The blue rectangle emphasizes the differences of two evolution process, which also represents the typical difference between \textbf{merge} and \textbf{after-merge} systems.}
    \label{merge-after-merge}
\end{figure}

A typical \textbf{merge} and an \textbf{after-merge} system are presneted in Figure \ref{merge-after-merge}. Notice that for the merge systems, the two stars come in contact and a common envelope process follows, similar to \textbf{spike} systems. Compared to the \textbf{spike} systems, we notice that the separation of the binary at the start of the common envelope is smaller in \textbf{merge} systems than in \textbf{spike} systems. This qualitatively explains the reason for the merging — the orbit does not have enough energy to eject the common envelope and two star merges. In contrast, the \textbf{after-merge} systems possesses more energy than the \textbf{merge} systems due to their larger intial separation. Hence, they either survives the common enevleope or do not come into contact at all. This explains why desired WD+MS systems form again as the initial separation $a_i$ increases.

This concludes our qualitative analysis on the evolution results of $7\Msun + 1\Msun$ through stable mass transfer. Again, as shown in Figure \ref{stable_hi}, for stable mass transfer, the observed minium separation of wide post-mass transfer WD+MS systems $0.15 \au$ can only be achived in the \textbf{spike} region. That is, in our results, all wide post-mass transfer systems are due to the strong wind mass loss from the primary that widens the orbit significantly.

Similarly, in \cite{yamaguchi_lo}, the author listed several arguments against the formation of wide post-mass transfer binaries through stable mass transfer. As shown in Figure \ref{theory-observed}, the observed wide post-mass transfer WD+MS binaries do not match the theoretical $M_{\mathrm{WD}} - P_{\mathrm{orb}}$ relation calculated in \cite{rappaport1995relation}, where the grey region is set to be a factor of 2.5 away from the calculated line.

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{theory-observed.png}
    \caption{$M_{\mathrm{WD}} - P_{\mathrm{orb}}$ relation of observed wide post-mass transfer binaries. Binaries going through stable mass transfer is expected to fall in the grey region, as calculated in \cite{rappaport1995relation}.}
    \label{theory-observed}
\end{figure}

\section{Population Simulation}
hello

\bibliographystyle{apalike}
\bibliography{reference.bib}

\end{document}

### 2024-06-25 summaries/june-summary.tex
add population plot to summary

@@
 \newcommand{\Erec}{E_{\mathrm{rec}}}
+\newcommand{\Porb}{P_{\mathrm{orb}}}
+\newcommand{\Mwd}{M_{\mathrm{WD}}}
 
@@
 Usually, the binding energy of the common envelope can be included in a structure parameter $\lambda$ such that
-\[
+
+\begin{equation}
 	\Ebind = \frac{G M_i M_{\mathrm{env}}}{\lambda R_i}.   
-\]
+	\label{ebind}
+\end{equation}
+
 This $\lambda$ is represented in COSMIC as the \verb|lambdaf| flag in \verb|BSEDict|. Hence, we want to calculate the effective $\lambda$ in COSMIC that matches these binding energy models in MESA.
@@
 Also, it is only in this region that the final separation goes larger than the observed minimum $0.15 \au$.
-Hence, in the next subsection, we focus on this sharp increase and try to expain it qualitatively.
+Hence, in the next subsection, we focus on this sharp increase and try to explain it qualitatively.
 
@@
 They represent typical "pre-spike" binaries and "spike" binaries.
-The blue rectangle marks out the common first phase of evolution in both systems — contact and common evnelope mass transfer.}
+The blue rectangle marks out the common first phase of evolution in both systems — contact and common envelope mass transfer.}
     \label{pre-spike-spike}
@@
 
-A typical \textbf{merge} and an \textbf{after-merge} system are presneted in Figure \ref{merge-after-merge}.
+A typical \textbf{merge} and an \textbf{after-merge} system are presented in Figure \ref{merge-after-merge}.
 Notice that for the merge systems, the two stars come in contact and a common envelope process follows, similar to \textbf{spike} systems.
@@
 This qualitatively explains the reason for the merging — the orbit does not have enough energy to eject the common envelope and two star merges.
-In contrast, the \textbf{after-merge} systems possesses more energy than the \textbf{merge} systems due to their larger intial separation.
-Hence, they either survives the common enevleope or do not come into contact at all.
+In contrast, the \textbf{after-merge} systems possesses more energy than the \textbf{merge} systems due to their larger initial separation.
+Hence, they either survives the common envelope or do not come into contact at all.
 This explains why desired WD+MS systems form again as the initial separation $a_i$ increases.
@@
 This concludes our qualitative analysis on the evolution results of $7\Msun + 1\Msun$ through stable mass transfer.
-Again, as shown in Figure \ref{stable_hi}, for stable mass transfer, the observed minium separation of wide post-mass transfer WD+MS systems $0.15 \au$ can only be achived in the \textbf{spike} region.
+Again, as shown in Figure \ref{stable_hi}, for stable mass transfer, the observed minimum separation of wide post-mass transfer WD+MS systems $0.15 \au$ can only be achieved in the \textbf{spike} region.
 That is, in our results, all wide post-mass transfer systems are due to the strong wind mass loss from the primary that widens the orbit significantly.
@@
 \section{Population Simulation}
-hello
+After investigating how the final separation depends on various parameters in individual binary star systems, we now shift our focus onto the distribution of final orbital period in a population of binaries (or equivalently final separation since $\Porb^2 \sim a^3$).
+
+\subsection{Methods}
+
+First of all, we sample the binary population using the independent sampler in COSMIC.
+We set the primary mass model to \verb|korupa01|, eccentricity model to \verb|uniform|, and orbital period model, together with the binary fraction model to \verb|moe19|.
+We sample a total number of about 10000 binaries with solar metallicity.
+The distribution of primary mass and the orbital period in the sampled population is shown in the Figure \ref{sample-distribution}.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\linewidth]{sample-distribution.png}
+	\caption{The distribution of primary mass, companion mass, and orbital period in the sampled population using independent sampler.
+\emph{Left:} Distribution of primary mass and companion mass in log scale.
+\emph{Right:} Distribution of orbital period $\Porb$ in log scale.}
+	\label{sample-distribution}
+\end{figure}
+
+Now to investigate how the final white dwarf mass and orbital period depends on \verb|lambdaf|, we simulation the sampled population with five different \verb|lambdaf| values, $-5$, $-15$, $-25$, $-35$, and $-45$, which corresponds to $\lambda = 5$, $15$, $25$, $35$ and $45$ in Equation \ref{ebind}.
+After that, for each binary star system in the population, we select the moment when CE is finished and when the system is a WD+MS one to record the white dwarf mass $\Mwd$ and final orbital period $P_{\mathrm{orb}, f}$.
+We also select the moment when the first RLOF started to record the initial orbital separation $P_{\mathrm{orb}, i}$.
+This will later on help us compare the difference and how much the orbit widens.
+
+\subsection{Results}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.5\linewidth]{Mwd-P scatter.png}
+	\caption{Scatter plot of white dwarf mass $\Mwd$ and final orbital period after common envelope in log scale $\log_{10}\Porb$.
+Different colors indicates results for different $\lambda$ values.
+For each different $\lambda$ value the same initial population is used.}
+	\label{Mwd-P-scatter}
+\end{figure}
+
+To show how the change in $\lambda$ affects the final orbital period, a scatter plot presenting the correlation of final orbital period $\Porb$ and white dwarf mass $\Mwd$ is presented in Figure \ref{Mwd-P-scatter}.
+We can see from the figure that when $\lambda$ increases, the final orbital period also increases.
+This is obvious for low $\Mwd$.
+This correlation is expected since for larger $\lambda$, less energy is require to unbind the common envelope.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\linewidth]{kstar-map-rearr.png}
+	\caption{Each panel: for this $\lambda$ value, scatter plot of white dwarf mass $\Mwd$ and final orbital period after common envelope in log scale $\log_{10}\Porb$.
+Different colors represent different kstar type of the primary at the start of mass transfer.}
+	\label{kstar-map}
+\end{figure}
+
+To also show how the final orbital separation $\Porb$ depends on the \verb|kstar_1| type at the start of first RLOF, we plot final orbital period $\Porb$ against white dwarf mass $\Mwd$ for each \verb|lambdaf| separately, and use different color to distinguish between different \verb|kstar_1| type at the start of mass transfer.
+The plot is shown in Figure \ref{kstar-map}.
+We can see that if the star is at later stages of evolution (\verb|kstar| large) when mass transfer starts, then the resulting white dwarf mass $\Mwd$ and orbital period $\Porb$ is also larger.
+Notice that few systems start the mass transfer at \verb|kstar_1 = 2| (Hertzsprung Gap) and \verb|kstar_1 = 4| (Core Helium Burning).
+This is because stars crosses Hertzsprung gap in a short timescale, leaving little time for interaction.
+Also, stellar radius is small when going through Core Helium Burning, reducing the possibility of interactions that starts when \verb|kstar_1 = 4|.
+
+To visualize the difference in final orbital period $\Porb$, and how they corresponds to different \verb|kstar_1| type at the start of RLOF, we create scatter plot in Figure \ref{dif-kstar map} of the difference in orbital period $\Delta P = P_{\mathrm{orb}}^{\lambda = 35} - P_{\mathrm{orb}}^{\lambda = 5}$ against the white dwarf mass $\Mwd$.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.5\linewidth]{dif-kstar map.png}
+	\caption{Scatter plot of difference in orbital period $\Delta P = P_{\mathrm{orb}}^{\lambda = 35} - P_{\mathrm{orb}}^{\lambda = 5}$ against the white dwarf mass $\Mwd$.
+Different colors indicates different kstar type of the primary at the start of mass transfer.}
+	\label{dif-kstar map}
+\end{figure}
+
+From Figure \ref{dif-kstar map}, we can see that for large \verb|kstar_1| type at the start of RLOF, $\Delta P = P_{\mathrm{orb}}^{\lambda = 35} - P_{\mathrm{orb}}^{\lambda = 5}$ is also larger.
+To explain this, we plot the how final white dwarf mass $\Mwd$, difference in final orbital period $\Delta \Porb$, and orbital period before mass transfer $P_{\mathrm{orb}, i}$ related to each other in Figure \ref{dif-kstar-exp}.
+We see that the larger initial orbital period $P_{\mathrm{orb}, i}$, the more significant effect does $\lambda$ have on the final orbital period.
+This is not unexpected since during mass transfer, the separation changes as $\dot{a} / a = \text{const}$.
+Hence $\dot{\Porb} / \Porb = \text{const}$.
+Therefore, the larger the initial orbital period, the larger it changes, and thus the larger effect does $\lambda$ have.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\linewidth]{dif-kstar two plot.png}
+	\caption{Left panel: scatter plot of initial orbital period at the start of RLOF $P_{\mathrm{orb}, i}$ against the final white dwarf mass $\MWD$.
+The color of the dots indicates how the final orbital period $P_{\mathrm{orb}, f}$ changes with respect to $\lambda=35$ and $\lambda=5$.
+That is, the color indicates the magnitude of $\Delta P = P_{\mathrm{orb}}^{\lambda = 35} - P_{\mathrm{orb}}^{\lambda = 5}$.
+Right panel: scatter plot of $\log_{10} \Delta \Porb$ against the initial orbital period in log scale $\log_{10} P_{\mathrm{orb}, i}$ at the start of RLOF.
+Different colors indicates different kstar type of the primary at the start of RLOF.}
+	\label{dif-kstar-exp}
+\end{figure}
 

### 2024-06-29 summaries/june-summary.tex
add mesa track comparison, modify summaries

@@
 \newcommand{\Porb}{P_{\mathrm{orb}}}
-\newcommand{\Mwd}{M_{\mathrm{WD}}}
-
 \newcommand{\au}{\, \mathrm{au}}
@@
 
+\section{Introduction}
+Common envelope (CE) is the outcome of unstable mass transfer.
+During CE, both stars orbit inside an envelope and spiral inward.
+If the energy liberated in this process is enough to eject the CE, then the result of CE process is a close binary.
+If, on the contrary, the liberated energy is not enough to eject the envelope, then the result of the process is a merger.
+In observation results, a number of WD+MS binaries are wider than expected, and CE is expected to be the main channel to form these white dwarf binaries \cite{yamaguchi_hi, yamaguchi_lo}.
+In the MESA model utilized by \cite{yamaguchi_hi, yamaguchi_lo}, wide post-CE WD+MS binaries can be formed in a certain range of initial separation with only gravitational and internal energy included in the calculation.
+
+In this paper, we are going to explore the formation of these wide post-mass transfer WD+MS binaries with COSMIC model.
+We perform COSMIC simulation on both individual binary star systems and sampled population to explore the formation of WD+MS binaries under different conditions.
+In section 2, we are going to investigate formation of individual wide WD+MS binaries through CE for a variety of variables, including initial mass, initial separation, common envelope efficiency, and the energy budget of the CE.
+In section 3, we are going to explore the formation of individual wide WD+MS binaries through stable mass transfer.
+In section 4, we will research on how different COSMIC model affect simulation results of a sample population of binary star systems, and how the formation of post WD+MS binaries depends on various parameters.
+
 \section{Common Envelope}
@@
 In \cite{yamaguchi_hi}, the lower limit of WD mass $M_{\mathrm{MD, min}}$ in the observational sample is calculated to be $1.244\Msun \sim 1.418\Msun$.
-The corresponding mass of MS progenitor is expected to be in the range of $6 \sim 9 \Msun$, and the companion mass has median around $1\Msun$ In \cite{yamaguchi_lo}, the WD mass in the Gaia sample ranges from $0.5\Msun$ to $0.8\Msun$, which corresponds to progenitor mass of $1 \Msun$ to $3 \Msun$.
+The corresponding mass of MS progenitor is expected to be in the range of $6 \sim 9 \Msun$, and the companion mass has median around $1\Msun$.
+In \cite{yamaguchi_lo}, the WD mass in the Gaia sample ranges from $0.5\Msun$ to $0.8\Msun$, which corresponds to progenitor mass of $1 \Msun$ to $3 \Msun$.
 The companions mass $\Mstar$ has a median of $0.85 \Msun$.
@@
 \end{figure}
-
+% =========================== Current Position ===========================
 \subsubsection{Discussions}
@@
 We notice that the first jump is probably because the evolution process changes from \textbf{3-7-8-4-3-7-8-4} to \textbf{3-7-8-7-8-4}.
-This means that the mass transfer changes from dynamically stable to dynamically unstable, leading to more mass loss during a shorter timescale and thus larger final separation.
+This means that for relatively larger initial separation $a_i$, the primary fills its Roche lobe all the time, leading to more mass loss during a shorter timescale and thus larger final separation.
 Also, the second jump is probably because \verb|kstar_1| changes from \verb|8| (Naked Helium Star Hertzsprung Gap) to \verb|9| (Naked Helium Star Giant Branch) at the start of the second common envelope.
-This changes the binding energy of the common evenlope and thus changes the final separation.
+This changes the binding energy of the common envelope and thus changes the final separation.
 
@@
 We use the following method to build the connection between COSMIC model and MESA model.
-For each fixed initial selaration $a_i$, we loop through the COSMIC results with the same initial separation, and find the \verb|lambdaf| value that results in a final separation closest to MESA model results, documented in the figures of \cite{yamaguchi_hi}.
+For each fixed initial separation $a_i$, we loop through the COSMIC results with the same initial separation, and find the \verb|lambdaf| value that results in a final separation closest to MESA model results, documented in the figures of \cite{yamaguchi_hi}.
 We record this \verb|lambdaf| as the effective $\lambda$ that matches the binding energy formalism at initial separation $a_i$.
 
-For $\Ebind = \Egrav + \Eint$, the fitting results is presented in Figure \ref{fit_hi}, where each line corresponds to different common envelope efficiencies $\alphace$ and differnt energy budgets (either $\Egrav$ only or $\Egrav + \Eint$).
+For $\Ebind = \Egrav + \Eint$, the fitting results is presented in Figure \ref{fit_hi}, where each line corresponds to different common envelope efficiencies $\alphace$ and different energy budgets (either $\Egrav$ only or $\Egrav + \Eint$).
 
@@
 
-To furthur compare with the default $\lambda$ value in COSMIC, we include in the right panel an extra line which represents the default $\lambda$ value in COSMIC.
+To further compare with the default $\lambda$ value in COSMIC, we include in the right panel an extra line which represents the default $\lambda$ value in COSMIC.
 This default $\lambda$ in the COSMIC model is calculated following Appendix A of \cite{claeys2014theoretical}.
@@
 The results is shown in Figure \ref{fit_cmp_hi}.
-Notice that the dafult $\lambda$ value is at order $\sim 1$, far smaller than the effective $\lambda$ needed to recreate results in \cite{yamaguchi_hi}.
+Notice that the default $\lambda$ value is at order $\sim 1$, far smaller than the effective $\lambda$ needed to recreate results in \cite{yamaguchi_hi}.
 
@@
 Now to investigate how the final white dwarf mass and orbital period depends on \verb|lambdaf|, we simulation the sampled population with five different \verb|lambdaf| values, $-5$, $-15$, $-25$, $-35$, and $-45$, which corresponds to $\lambda = 5$, $15$, $25$, $35$ and $45$ in Equation \ref{ebind}.
-After that, for each binary star system in the population, we select the moment when CE is finished and when the system is a WD+MS one to record the white dwarf mass $\Mwd$ and final orbital period $P_{\mathrm{orb}, f}$.
+After that, for each binary star system in the population, we select the moment when CE is finished and when the system is a WD+MS one to record the white dwarf mass $\MWD$ and final orbital period $P_{\mathrm{orb}, f}$.
 We also select the moment when the first RLOF started to record the initial orbital separation $P_{\mathrm{orb}, i}$.
@@
 	\includegraphics[width=0.5\linewidth]{Mwd-P scatter.png}
-	\caption{Scatter plot of white dwarf mass $\Mwd$ and final orbital period after common envelope in log scale $\log_{10}\Porb$.
+	\caption{Scatter plot of white dwarf mass $\MWD$ and final orbital period after common envelope in log scale $\log_{10}\Porb$.
 Different colors indicates results for different $\lambda$ values.
@@
 
-To show how the change in $\lambda$ affects the final orbital period, a scatter plot presenting the correlation of final orbital period $\Porb$ and white dwarf mass $\Mwd$ is presented in Figure \ref{Mwd-P-scatter}.
+To show how the change in $\lambda$ affects the final orbital period, a scatter plot presenting the correlation of final orbital period $\Porb$ and white dwarf mass $\MWD$ is presented in Figure \ref{Mwd-P-scatter}.
 We can see from the figure that when $\lambda$ increases, the final orbital period also increases.
-This is obvious for low $\Mwd$.
+This is obvious for low $\MWD$.
 This correlation is expected since for larger $\lambda$, less energy is require to unbind the common envelope.
@@
 	\includegraphics[width=\linewidth]{kstar-map-rearr.png}
-	\caption{Each panel: for this $\lambda$ value, scatter plot of white dwarf mass $\Mwd$ and final orbital period after common envelope in log scale $\log_{10}\Porb$.
+	\caption{Each panel: for this $\lambda$ value, scatter plot of white dwarf mass $\MWD$ and final orbital period after common envelope in log scale $\log_{10}\Porb$.
 Different colors represent different kstar type of the primary at the start of mass transfer.}
@@
 
-To also show how the final orbital separation $\Porb$ depends on the \verb|kstar_1| type at the start of first RLOF, we plot final orbital period $\Porb$ against white dwarf mass $\Mwd$ for each \verb|lambdaf| separately, and use different color to distinguish between different \verb|kstar_1| type at the start of mass transfer.
+To also show how the final orbital separation $\Porb$ depends on the \verb|kstar_1| type at the start of first RLOF, we plot final orbital period $\Porb$ against white dwarf mass $\MWD$ for each \verb|lambdaf| separately, and use different color to distinguish between different \verb|kstar_1| type at the start of mass transfer.
 The plot is shown in Figure \ref{kstar-map}.
-We can see that if the star is at later stages of evolution (\verb|kstar| large) when mass transfer starts, then the resulting white dwarf mass $\Mwd$ and orbital period $\Porb$ is also larger.
+We can see that if the star is at later stages of evolution (\verb|kstar| large) when mass transfer starts, then the resulting white dwarf mass $\MWD$ and orbital period $\Porb$ is also larger.
 Notice that few systems start the mass transfer at \verb|kstar_1 = 2| (Hertzsprung Gap) and \verb|kstar_1 = 4| (Core Helium Burning).
@@
 
-To visualize the difference in final orbital period $\Porb$, and how they corresponds to different \verb|kstar_1| type at the start of RLOF, we create scatter plot in Figure \ref{dif-kstar map} of the difference in orbital period $\Delta P = P_{\mathrm{orb}}^{\lambda = 35} - P_{\mathrm{orb}}^{\lambda = 5}$ against the white dwarf mass $\Mwd$.
+To visualize the difference in final orbital period $\Porb$, and how they corresponds to different \verb|kstar_1| type at the start of RLOF, we create scatter plot in Figure \ref{dif-kstar map} of the difference in orbital period $\Delta P = P_{\mathrm{orb}}^{\lambda = 35} - P_{\mathrm{orb}}^{\lambda = 5}$ against the white dwarf mass $\MWD$.
 
@@
 	\includegraphics[width=0.5\linewidth]{dif-kstar map.png}
-	\caption{Scatter plot of difference in orbital period $\Delta P = P_{\mathrm{orb}}^{\lambda = 35} - P_{\mathrm{orb}}^{\lambda = 5}$ against the white dwarf mass $\Mwd$.
+	\caption{Scatter plot of difference in orbital period $\Delta P = P_{\mathrm{orb}}^{\lambda = 35} - P_{\mathrm{orb}}^{\lambda = 5}$ against the white dwarf mass $\MWD$.
 Different colors indicates different kstar type of the primary at the start of mass transfer.}
@@
 From Figure \ref{dif-kstar map}, we can see that for large \verb|kstar_1| type at the start of RLOF, $\Delta P = P_{\mathrm{orb}}^{\lambda = 35} - P_{\mathrm{orb}}^{\lambda = 5}$ is also larger.
-To explain this, we plot the how final white dwarf mass $\Mwd$, difference in final orbital period $\Delta \Porb$, and orbital period before mass transfer $P_{\mathrm{orb}, i}$ related to each other in Figure \ref{dif-kstar-exp}.
+To explain this, we plot the how final white dwarf mass $\MWD$, difference in final orbital period $\Delta \Porb$, and orbital period before mass transfer $P_{\mathrm{orb}, i}$ related to each other in Figure \ref{dif-kstar-exp}.
 We see that the larger initial orbital period $P_{\mathrm{orb}, i}$, the more significant effect does $\lambda$ have on the final orbital period.

### 2024-06-29 summaries/paper-draft.tex
add mesa track comparison, modify summaries

\documentclass[linenumbers]{aastex631}

\newcommand{\vdag}{(v)^\dagger}
\newcommand\aastex{AAS\TeX}
\newcommand\latex{La\TeX}
\newcommand{\Msun}{\mathrm{M_{\odot}}}
\newcommand{\Rsun}{\mathrm{R_{\odot}}}
\newcommand{\MWD}{M_{\mathrm{WD}}}
\newcommand{\Mstar}{M_{\star}}
\newcommand{\alphace}{\alpha_{\mathrm{CE}}}
\newcommand{\alphath}{\alpha_{\mathrm{th}}}
\newcommand{\alpharec}{\alpha_{\mathrm{rec}}}
\newcommand{\Ebind}{E_{\mathrm{bind}}}
\newcommand{\Egrav}{E_{\mathrm{grav}}}
\newcommand{\Eint}{E_{\mathrm{int}}}
\newcommand{\Eth}{E_{\mathrm{th}}}
\newcommand{\Erec}{E_{\mathrm{rec}}}
\newcommand{\Porb}{P_{\mathrm{orb}}}
\newcommand{\au}{\, \mathrm{au}}

\begin{document}

\title{Wide Post-Mass Transfer White Dwarf Binaries}

%\author[0000-0002-0786-7307]{Greg J. Schwarz}
%\affiliation{American Astronomical Society \\
%1667 K Street NW, Suite 800 \\
%Washington, DC 20006, USA}

\author{Runqiu Ye}
\affiliation{Carnegie Mellon University}

%\collaboration{20}{(AAS Journals Data Editors)}
%
%\author{F.X Timmes}
%\affiliation{Arizona State University}
%\affiliation{AAS Journals Associate Editor-in-Chief}
%
%\author{Amy Hendrickson}
%\altaffiliation{AASTeX v6+ programmer}
%\affiliation{TeXnology Inc.}
%
%\author{Julie Steffen}
%\affiliation{AAS Director of Publishing}
%\affiliation{American Astronomical Society \\
%1667 K Street NW, Suite 800 \\
%Washington, DC 20006, USA}

\begin{abstract}

(***TO-DO***) This example manuscript is intended to serve as a tutorial and template for
authors to use when writing their own AAS Journal articles. The manuscript
includes a history of \aastex\ and includes figure and table examples to illustrate these features. Information on features not explicitly mentioned in the article can be viewed in the manuscript comments or more extensive online
documentation. Authors are welcome replace the text, tables, figures, and
bibliography with their own and submit the resulting manuscript to the AAS
Journals peer review system.  The first lesson in the tutorial is to remind
authors that the AAS Journals, the Astrophysical Journal (ApJ), the
Astrophysical Journal Letters (ApJL), the Astronomical Journal (AJ), and
the Planetary Science Journal (PSJ) all have a 250 word limit for the 
abstract\footnote{Abstracts for Research Notes of the American Astronomical 
Society (RNAAS) are limited to 150 words}.  If you exceed this length the
Editorial office will ask you to shorten it. This abstract has 161 words.

\end{abstract}

\keywords{(***TO-DO***) Classical Novae (251) --- Ultraviolet astronomy(1736) --- History of astronomy(1868) --- Interdisciplinary astronomy(804)}

\section{Introduction} \label{sec:intro}

Common envelope (CE) is the outcome of unstable mass transfer. During CE, both stars orbit inside an envelope and spiral inward. If the energy liberated in this process is enough to eject the CE, then the result of CE process is a close binary. If, on the contrary, the liberated energy is not enough to eject the envelope, then the result of the process is a merger. In observation results, a number of WD+MS binaries are wider than expected, and CE is expected to be the main channel to form these white dwarf binaries \cite{yamaguchi_hi, yamaguchi_lo}. In the MESA model utilized by \cite{yamaguchi_hi, yamaguchi_lo}, wide post-CE WD+MS binaries can be formed in a certain range of initial separation with only gravitational and internal energy included in the calculation.

In this paper, we are going to explore the formation of these wide post-mass transfer WD+MS binaries with COSMIC model. We perform COSMIC simulation on both individual binary star systems and sampled population to explore the formation of WD+MS binaries under different conditions. In section \ref{sec:ce}, we are going to investigate formation of individual wide WD+MS binaries through CE for a variety of variables, including initial mass, initial separation, CE efficiency, and the energy budget of the CE. In section \ref{sec:stable}, we are going to explore the formation of individual wide WD+MS binaries through stable mass transfer. In section \ref{sec:population}, we will research on how different COSMIC model affect simulation results of a sample population of binary star systems, and how the formation of post WD+MS binaries depends on various parameters.

\section{Common Envelope} \label{sec:ce}
In \cite{yamaguchi_hi, yamaguchi_lo}, the author discussed the formation of WD+MS binaries through CE for both high-mass and low-mass systems, and it was found that wide WD+MS systems can be formed for certain energy budgets and certain initial separation in the MESA model. We hope to explore similar models in COSMIC.

In \cite{yamaguchi_hi, yamaguchi_lo}, to explore the dependency of the final separation $a_f$ in the post-CE binary on the initial separation $a_i$ in the pre-MT binary, the author first run a stellar evolution model of the primary star using MESA. Its evolution up to AGB phase is followed. With the model, the binding energy of the CE is calculated as

\begin{equation}
	E_{\mathrm{bind}} = E_{\mathrm{grav}} + E_{\mathrm{int}} = \int_{M_{\mathrm{core}}}^{M_{\mathrm{tot}}} -\frac{Gm}{r(m)} + U(m) dm,
	\label{budget-int}
\end{equation}
or
\begin{equation}
	E_{\mathrm{bind}} = E_{\mathrm{grav}} + \alphath E_{\mathrm{th}} + \alpharec E_{\mathrm{rec}}.
	\label{budget-th-rec}
\end{equation}
After that, the change of separation through CE evolution is calculated using
\begin{equation}
	E_{\mathrm{bind}} = \alphace \left(-\frac{G\MWD \Mstar}{2a_f}+\frac{GM_i \Mstar}{2a_i}\right).
	\label{ebind-sep}
\end{equation}

In COSMIC, the CE is modeled using a structural parameter $\lambda$, where
\begin{equation}
	\Ebind = \frac{G M_i M_{\mathrm{env}}}{\lambda R_i}.   
	\label{ebind}
\end{equation}
This $\lambda$ is represented in COSMIC as the \verb|lambdaf| flag in \verb|BSEDict|. To compare MESA model with COSMIC, we intend to compare how the energy budget utilized by \cite{yamaguchi_hi, yamaguchi_lo} matches with the CE model in COSMIC. We approach this by calculating the effective $\lambda$ value for each of the energy budget used in \cite{yamaguchi_hi, yamaguchi_lo}.

\subsection{High Mass Systems $(7 \Msun + 1 \Msun)$} \label{subsec:high}

\subsubsection{Methods and results}

In \cite{yamaguchi_hi}, the lower limit of WD mass $M_{\mathrm{MD, min}}$ in the observational sample is calculated to be $1.244\Msun \sim 1.418\Msun$. The corresponding mass of MS progenitor is expected to be in the range of $6 \sim 9 \Msun$, and the companion mass has median around $1\Msun$. MESA results for $7\Msun + 1\Msun$ systems is documented in \cite{yamaguchi_hi}. Hence, we also consider a $7\Msun + 1\Msun$ evolution model in COSMIC and compare the results with those in \cite{yamaguchi_hi}.

We run COSMIC binary evolution simulation of a $7\Msun + 1\Msun$ system and investigate how the final separation $a_f$ depends on the initial separation $a_i$ and the CE parameter $\lambda$. We vary the initial separation in range $2 \sim 8 \au$ with $400$ steps. We vary the CE flag \verb|lambdaf| in range $0 \sim -100$, which corresponds to $\lambda = 0 \sim 100$, with $2000$ steps. Also, we run these parameters with four different CE efficiencies — $\alphace = 1$, $\alphace = 0.9$, $\alphace = 0.6$, and $\alphace = 0.3$, the same as documented in \cite{yamaguchi_hi}. In total, for $4$ different CE efficiency $\alphace$, we run a grid of $80000$ binary star systems with different initial separation $a_i$ and CE parameter $\lambda$.

After simulation is finished, we select the time-step when the system finished CE for the first time (\verb|evol_type = 8|). At this moment, if the system is a WD+MS system, we record their separation as the final separation $a_f$. Otherwise, we set $a_f = 0$ and consider this initial condition unable to produce a WD+MS system.

We create four heat maps for the four different CE efficiencies. In Figure \ref{res_hi}, we plot the final separation $a_f$ of the WD+MS systems against initial separation $a_i$ and the CE flag \verb|lambdaf|$= - \lambda$.

\begin{figure}
	\centering
	\includegraphics[width=0.8\linewidth]{7+1results.png}
	\caption{ Heat map of final separation $a_f$ against initial separation $a_i$ and CE flag lambdaf$=-\lambda$. Different panels represent four different CE efficiencies $\alphace = 1$, $0.9$, $0.6$, and $0.3$. The white and yellow contour corresponds to $a_f = 0.01 \au$ and $a_f = 0.15 \au$ respectively. WD+MS binaries is not possible for separation smaller than $0.01 \au$, and $0.15 \au$ is the minimum separation of wide WD+MS binaries in observational results. Outside the white contour no desired system forms.}
	\label{res_hi}
\end{figure}

\subsubsection{Discussions}

There are several interesting points of the results. First of all, as $\alphace$ decreases, the final separation decreases in general. This can be explained by Equation \ref{ebind-sep}. Qualitatively, note that more energy from the orbit is needed to eject the envelope for low $\alphace$, so the resulting $a_f$ is expected to be smaller.

It is worth noting that most of these systems experience two mass transfer phases in COSMIC before reaching WD+MS. That is, during the evolution process, two CE processes take place, between which there is a period of time. By checking the \verb|bpp| array, which documents the evolution processes of the binary star systems, we found that (***TO-DO***) fraction of all our simulated binaries experience two CE processes before reaching WD+MS. When the primary loses too much mass during the first CE, the CE process ends. After the primary fills its Roche lobe and has a large enough envelope again, the CE process resumes and causes two CE processes in total.

From Figure \ref{res_hi}, we also notice that the final separation $a_f$ jumps obviously at about $2.9 \au$, and $7.1 \au$. After investigating the \verb|bpp| array of these simulations, we find out the jump at about $2.9 \au$ and $7.1 \au$ is likely due to the difference of \verb|kstar_1| at the start of RLOF. For $a_i < 2.9 \au$, RLOF starts when the primary is on the giant branch (GB) with \verb|kstar_1 3|. For $a_i = 2.9 \au$, RLOF starts when the primary is experiencing core Helium burning (CHeB) with \verb|kstar_1 = 4|. For $2.9 \au < a_i < 7.1 \au$, RLOF starts when the primary is on E-AGB with \verb|kstar_1 = 5|. For $a_i > 7.1 \au$, RLOF starts when the primary is on TP-AGB with \verb|kstar_1 = 6|. For different star type, the structure and mass of the envelope is different, leading to different binding energy $\Ebind$ and thus different final separation $a_f$.

\begin{figure}
	\centering
	\includegraphics[width = 0.5\linewidth]{jump-zoom.png}
	\caption{Dependence of final separation $a_f$ on initial separation $a_i$ for $\alphace = 0.3$. Simulation results of $\lambda_f = 60$, $70$, and $80$ are shown. The red circles emphasizes the jump of final separation as initial separation increases.}
	\label{jump-zoom}
\end{figure}

Inspecting more closely, we find that the final separation $a_f$ jumps at about $5.2 \sim 5.4 \au$. This is especially obvious with $\alphace = 0.3$. To illustrate, we zoom in the simulation results of $\alphace = 0.3$, and present the change of final separation $a_f$ as initial separation $a_i$ increases in Figure \ref{jump-zoom}. From here, we can clearly see two jumps — one between $5.1\au$ and $5.2\au$, the other at $5.34\au$. To explain this, we again go to the \verb|bpp| array of these systems. We notice that the first jump is probably because the evolution process changes from \textbf{3-7-8-4-3-7-8-4} to \textbf{3-7-8-7-8-4}. This means that for relatively larger initial separation $a_i$, the primary fills its Roche lobe all the time, leading to more mass loss during a shorter timescale and thus larger final separation. Also, the second jump is probably because \verb|kstar_1| changes from \verb|8| (Naked Helium Star Hertzsprung Gap) to \verb|9| (Naked Helium Star Giant Branch) at the start of the second common envelope. This changes the binding energy of the common envelope and thus changes the final separation.

Finally, notice that there exists a small triangular gap in the figure for $\alphace = 1$. This is because \verb|kstar_1| becomes a neutron star and no rows in the \verb|bpp| gets selected.

In conclusion, with a large \verb|lambdaf|, it is possible to create wide post-mass transfer WD+MS systems with final separation $a_f > 0.15 \au$, which is the minimal separation of our observed objects. In next section, we will compare these results with default settings in COSMIC and with results in \cite{yamaguchi_hi}, in order to investigate the possibility of forming wide WD+MS systems through CE.

\subsubsection{Comparison}
After we have obtained the simulation results for different initial conditions, now we can calculate the effective $\lambda$ that matches with results in \cite{yamaguchi_hi}. For each fixed initial separation $a_i$, we loop through the COSMIC results with the same initial separation, and find the \verb|lambdaf| value that results in a final separation closest to MESA model results. We record this \verb|lambdaf| as the effective $\lambda$ that matches the binding energy formalism at initial separation $a_i$.

\begin{figure}
	\centering
	\includegraphics[width=0.9\linewidth]{7+1fit.png}
	\caption{Effective $\lambda$ fitting results of a $7\Msun + 1\Msun$ system for common envelope structure in MESA. Left panel: MESA results in \cite{yamaguchi_hi} of how final separation $a_f$ depends on initial separation $a_i$. Central panel: COSMIC results of how final separation $a_f$ depends on initial separation $a_i$, produced by the effective $\lambda$ calculated. Right panel: effective $\lambda$ in COSMIC that corresponds to the MESA results for each initial separation $a_i$. The red dashed line in the left and middle panel represents $0.15 \au$, which is the minimal separation for observed wide post-CE WD+MS binaries.}
	\label{fit_hi}
\end{figure}

We present the fitting results for energy budget $\Ebind = \Egrav + \Eint$ in Figure \ref{fit_hi}. Notice that in the central panel, there is a small bump of the blue line. This is because at $6 \sim 7 \au$, we cannot reach such low final separation in COSMIC while forming a WD+MS binary at the same time. From the left panel, we can see that it is possible to create WD+MS binaries with final separation $>0.15 \au$ in MESA as long as internal energy $\Eint$ is included. However, in the right panel we show that this correspond to very large \verb|lambdaf| in COSMIC ($\lambda > \sim 20$).

\begin{figure}
	\centering
	\includegraphics[width=0.9\linewidth]{7+1cmp.png}
	\caption{Same as Figure \ref{fit_hi}, but with default $\lambda$ in COSMIC model included in the right panel.}
	\label{fit_cmp_hi}
\end{figure}

To further compare with the default $\lambda$ value in COSMIC, we include in the right panel an extra line which represents the default $\lambda$ value in COSMIC. This default $\lambda$ in the COSMIC model is calculated following Appendix A of \cite{claeys2014theoretical}. For all of our case, we have $M_{\mathrm{env}} > 1$. The results is shown in Figure \ref{fit_cmp_hi}. Notice that the default $\lambda$ value is at order $\sim 1$, far smaller than the effective $\lambda$ needed to recreate results in \cite{yamaguchi_hi}.

\begin{figure}
	\centering
	\includegraphics[width=0.9\linewidth]{7+1ebcmp.png}
	\caption{Same as Figure \ref{fit_cmp_hi}, but with a different energy budget — $\Ebind = \Egrav + \alphath \Eth + \alpharec \Erec$.}
	\label{fit_cmp_eb_hi}
\end{figure}

In Figure \ref{fit_cmp_eb_hi}, we present the same results for another energy budget of the CE process — $\Ebind = \Egrav + \alphath \Eth + \alpharec \Erec$.

\subsection{Low Mass Systems $(1.5 \Msun + 0.85 \Msun)$} \label{subsec:low}
In \cite{yamaguchi_lo}, the WD mass in the Gaia sample ranges from $0.5\Msun$ to $0.8\Msun$, which corresponds to progenitor mass of $1 \Msun$ to $3 \Msun$. The companions mass has a median of $0.85 \Msun$. MESA results for $1.5\Msun + 0.85\Msun$ systems is documented in \cite{yamaguchi_hi}. Hence, we consider a $1.5\Msun + 0.85\Msun$ evolution model in COSMIC and compare the results with those in \cite{yamaguchi_hi}.

\section{Stable Mass Transfer} \label{sec:stable}

\section{Population} \label{sec:population}

\bibliography{reference.bib}
\bibliographystyle{aasjournal}

\end{document}

### 2024-07-08 summaries/june-summary.tex
reformatted paper draft, continue

@@
     \centering
-    \includegraphics[width=\linewidth]{yamaguchi-7+1.png}
+    \includegraphics[width=\linewidth]{fig/yamaguchi-7+1.png}
     \caption{Results of how final separation $a_f$ after common envelope depends on initial separation $a_i$ for different common envelope efficiencies and energy budgets in \cite{yamaguchi_hi}.}
@@
     \centering
-    \includegraphics[width=0.9\linewidth]{7+1results.png}
+    \includegraphics[width=0.9\linewidth]{fig/7+1results.png}
     \caption{COSMIC results of how final separation $a_f$ depends on initial separation $a_i$ and common envelope flag "lambdaf" for different common envelope efficiencies $\alphace$, where the black and red contour corresponds to $0.01 \au$ and $0.15 \au$ as in Figure \ref{yam_hi} respectively.
@@
     \centering
-    \includegraphics[width = 0.6\linewidth]{jump-zoom.png}
+    \includegraphics[width = 0.6\linewidth]{fig/jump-zoom.png}
     \caption{Dependence of final separation $a_f$ on initial separation $a_i$ for $\alphace = 0.3$.
@@
     \centering
-    \includegraphics[width=1\linewidth]{7+1fit.png}
+    \includegraphics[width=1\linewidth]{fig/7+1fit.png}
     \caption{Effective $\lambda$ fitting results of a $7\Msun + 1\Msun$ system for common envelope structure in MESA.
@@
     \centering
-    \includegraphics[width=\linewidth]{7+1cmp.png}
+    \includegraphics[width=\linewidth]{fig/7+1cmp.png}
     \caption{Same as Figure \ref{fit_hi}, but with default $\lambda$ in COSMIC model included in the right panel.}
@@
     \centering
-    \includegraphics[width=\linewidth]{7+1ebcmp.png}
+    \includegraphics[width=\linewidth]{fig/7+1ebcmp.png}
     \caption{Same as Figure \ref{fit_cmp_hi}, but with a different energy budget.}
@@
     \centering
-    \includegraphics[width=\linewidth]{yamaguchi-1.5+0.8.png}
+    \includegraphics[width=\linewidth]{fig/yamaguchi-1.5+0.8.png}
     \caption{Results of how final separation $a_f$ after common envelope depends on initial separation $a_i$ for different common envelope efficiencies and energy budgets in \cite{yamaguchi_lo}.
@@
     \centering
-    \includegraphics[width=0.9\linewidth]{1.5+0.8results.png}
+    \includegraphics[width=0.9\linewidth]{fig/1.5+0.8results.png}
     \caption{COSMIC results of how final separation $a_f$ depends on initial separation $a_i$ and common envelope flag "lambdaf" for different common envelope efficiencies $\alphace$, where the black and red contour corresponds to $0.01 \au$, $0.15 \au$, and $0.7 \au$ as in Figure \ref{yam_lo} respectively.
@@
     \centering
-    \includegraphics[width=\linewidth]{1.5+0.8cmp_whole.png}
+    \includegraphics[width=\linewidth]{fig/1.5+0.8cmp_whole.png}
     \caption{Effective $\lambda$ fitting results of a $1.5\Msun + 0.85\Msun$ system for common envelope structure in MESA.
@@
     \centering
-    \includegraphics[width=\linewidth]{1.5+0.8cmp_pt.png}
+    \includegraphics[width=\linewidth]{fig/1.5+0.8cmp_pt.png}
     \caption{Same as Figure \ref{fit_cmp_lo_pt} but only binding enregy of the \textbf{outer envelope} is included.}
@@
     \centering
-    \includegraphics[width=0.75\linewidth]{stable_hi_log.png}
+    \includegraphics[width=0.75\linewidth]{fig/stable_hi_log.png}
     \caption{COSMIC results of WD+MS binaries after stable mass transfer for $7\Msun + 1\Msun$ systems, with accretion efficiency $\beta = 0$, $\beta = 0.5$, and $\beta = 1$ respectively.
@@
     \centering
-    \includegraphics[width = 0.75\linewidth]{stable7+1ex.png}
+    \includegraphics[width = 0.75\linewidth]{fig/stable7+1ex.png}
     \caption{COSMIC results of WD+MS binaries after stable mass transfer for two systems — $7\Msun + 0.7\Msun$ with accretion efficiency $\beta = 0.1$, and $7\Msun + 0.5\Msun$ with accretion efficiency $\beta = 0.2$.
@@
     \centering
-    \includegraphics[width=\linewidth]{pre-spike-spike.png}
+    \includegraphics[width=\linewidth]{fig/pre-spike-spike.png}
     \caption{COSMIC simulation data of $7 \Msun + 1 \Msun$ systems with accretion efficiency $\beta = 0$ and initial separation $2.30 \au$ and $2.90 \au$ respectively.
@@
     \centering
-    \includegraphics[width=\linewidth]{sharp-inc-bpp.png}
+    \includegraphics[width=\linewidth]{fig/sharp-inc-bpp.png}
     \caption{COSMIC simulation of systems $7 \Msun + 1 \Msun$ with accretion efficiency $\beta = 0$ and initial separation $2.80 \au$ and $2.90 \au$ respectively.
@@
     \centering
-    \includegraphics[width=\linewidth]{merge-after-merge.png}
+    \includegraphics[width=\linewidth]{fig/merge-after-merge.png}
     \caption{COSMIC simulation of systems $7 \Msun + 1 \Msun$ with accretion efficiency $\beta = 0$ and initial separation $3.25 \au$ and $5.00 \au$ respectively.
@@
     \centering
-    \includegraphics[width=\linewidth]{theory-observed.png}
+    \includegraphics[width=\linewidth]{fig/theory-observed.png}
     \caption{$M_{\mathrm{WD}} - P_{\mathrm{orb}}$ relation of observed wide post-mass transfer binaries.
@@
 	\centering
-	\includegraphics[width=\linewidth]{sample-distribution.png}
+	\includegraphics[width=\linewidth]{fig/sample-distribution.png}
 	\caption{The distribution of primary mass, companion mass, and orbital period in the sampled population using independent sampler.
@@
 	\centering
-	\includegraphics[width=0.5\linewidth]{Mwd-P scatter.png}
+	\includegraphics[width=0.5\linewidth]{fig/Mwd-P scatter.png}
 	\caption{Scatter plot of white dwarf mass $\MWD$ and final orbital period after common envelope in log scale $\log_{10}\Porb$.
@@
 	\centering
-	\includegraphics[width=\linewidth]{kstar-map-rearr.png}
+	\includegraphics[width=\linewidth]{fig/kstar-map-rearr.png}
 	\caption{Each panel: for this $\lambda$ value, scatter plot of white dwarf mass $\MWD$ and final orbital period after common envelope in log scale $\log_{10}\Porb$.
@@
 	\centering
-	\includegraphics[width=0.5\linewidth]{dif-kstar map.png}
+	\includegraphics[width=0.5\linewidth]{fig/dif-kstar map.png}
 	\caption{Scatter plot of difference in orbital period $\Delta P = P_{\mathrm{orb}}^{\lambda = 35} - P_{\mathrm{orb}}^{\lambda = 5}$ against the white dwarf mass $\MWD$.
@@
 	\centering
-	\includegraphics[width=\linewidth]{dif-kstar two plot.png}
+	\includegraphics[width=\linewidth]{fig/dif-kstar two plot.png}
 	\caption{Left panel: scatter plot of initial orbital period at the start of RLOF $P_{\mathrm{orb}, i}$ against the final white dwarf mass $\MWD$.

### 2024-07-08 summaries/paper-draft-plain.tex
reformatted paper draft, continue

--- 2024-06-29 summaries/paper-draft.tex
@@
-\documentclass[linenumbers]{aastex631}
+\documentclass[12pt]{article}
+\usepackage[utf8]{inputenc}
+\usepackage[T1]{fontenc}
+\usepackage[top=0.60in, bottom=0.80in, left=0.60in, right=0.60in]{geometry}
+\usepackage{graphicx}
+\usepackage{siunitx}
+\usepackage{listings}
 
-\newcommand{\vdag}{(v)^\dagger}
-\newcommand\aastex{AAS\TeX}
-\newcommand\latex{La\TeX}
 \newcommand{\Msun}{\mathrm{M_{\odot}}}
@@
 
+\title{Wide Post-Mass Transfer White-Dwarf Binaries}
+\author{Runqiu Ye}
+\date{June, 2024}
+
 \begin{document}
 
-\title{Wide Post-Mass Transfer White Dwarf Binaries}
-
-%\author[0000-0002-0786-7307]{Greg J. Schwarz}
-%\affiliation{American Astronomical Society \\
-%1667 K Street NW, Suite 800 \\
-%Washington, DC 20006, USA}
-
-\author{Runqiu Ye}
-\affiliation{Carnegie Mellon University}
-
-%\collaboration{20}{(AAS Journals Data Editors)}
-%
-%\author{F.X Timmes}
-%\affiliation{Arizona State University}
-%\affiliation{AAS Journals Associate Editor-in-Chief}
-%
-%\author{Amy Hendrickson}
-%\altaffiliation{AASTeX v6+ programmer}
-%\affiliation{TeXnology Inc.}
-%
-%\author{Julie Steffen}
-%\affiliation{AAS Director of Publishing}
-%\affiliation{American Astronomical Society \\
-%1667 K Street NW, Suite 800 \\
-%Washington, DC 20006, USA}
-
-\begin{abstract}
-
-(***TO-DO***) This example manuscript is intended to serve as a tutorial and template for
-authors to use when writing their own AAS Journal articles. The manuscript
-includes a history of \aastex\ and includes figure and table examples to illustrate these features.
-Information on features not explicitly mentioned in the article can be viewed in the manuscript comments or more extensive online
-documentation. Authors are welcome replace the text, tables, figures, and
-bibliography with their own and submit the resulting manuscript to the AAS
-Journals peer review system.  The first lesson in the tutorial is to remind
-authors that the AAS Journals, the Astrophysical Journal (ApJ), the
-Astrophysical Journal Letters (ApJL), the Astronomical Journal (AJ), and
-the Planetary Science Journal (PSJ) all have a 250 word limit for the 
-abstract\footnote{Abstracts for Research Notes of the American Astronomical 
-Society (RNAAS) are limited to 150 words}.  If you exceed this length the
-Editorial office will ask you to shorten it. This abstract has 161 words.
-
-\end{abstract}
-
-\keywords{(***TO-DO***) Classical Novae (251) --- Ultraviolet astronomy(1736) --- History of astronomy(1868) --- Interdisciplinary astronomy(804)}
+\maketitle
 
@@
   \centering
-	\includegraphics[width=0.8\linewidth]{7+1results.png}
+  \includegraphics[width=0.8\linewidth]{fig/7+1results.png}
   \caption{ Heat map of final separation $a_f$ against initial separation $a_i$ and CE flag lambdaf$=-\lambda$.
@@
   \centering
-	\includegraphics[width = 0.5\linewidth]{jump-zoom.png}
+  \includegraphics[width = 0.5\linewidth]{fig/jump-zoom.png}
   \caption{Dependence of final separation $a_f$ on initial separation $a_i$ for $\alphace = 0.3$.
@@
   \centering
-	\includegraphics[width=0.9\linewidth]{7+1fit.png}
+  \includegraphics[width=0.9\linewidth]{fig/7+1fit.png}
   \caption{Effective $\lambda$ fitting results of a $7\Msun + 1\Msun$ system for common envelope structure in MESA.
@@
   \centering
-	\includegraphics[width=0.9\linewidth]{7+1cmp.png}
+  \includegraphics[width=0.9\linewidth]{fig/7+1cmp.png}
   \caption{Same as Figure \ref{fit_hi}, but with default $\lambda$ in COSMIC model included in the right panel.}
@@
   \centering
-	\includegraphics[width=0.9\linewidth]{7+1ebcmp.png}
+  \includegraphics[width=0.9\linewidth]{fig/7+1ebcmp.png}
   \caption{Same as Figure \ref{fit_cmp_hi}, but with a different energy budget — $\Ebind = \Egrav + \alphath \Eth + \alpharec \Erec$.}
@@
 \bibliography{reference.bib}
-\bibliographystyle{aasjournal}
+\bibliographystyle{apalike}
 

### 2024-07-08 summaries/paper-draft-plain.tex
modified high mass unstable part of summary and reran relevant code

@@
 \usepackage{listings}
+\usepackage[labelfont=bf]{caption}
 
-\newcommand{\Msun}{\mathrm{M_{\odot}}}
-\newcommand{\Rsun}{\mathrm{R_{\odot}}}
+\newcommand{\Msun}{\,\mathrm{M_{\odot}}}
+\newcommand{\Rsun}{\,\mathrm{R_{\odot}}}
 \newcommand{\MWD}{M_{\mathrm{WD}}}
@@
 
-In this paper, we are going to explore the formation of these wide post-mass transfer WD+MS binaries with COSMIC model.
+In this paper, we are going to explore the formation possibility of these wide post-mass transfer WD+MS binaries with COSMIC model.
 We perform COSMIC simulation on both individual binary star systems and sampled population to explore the formation of WD+MS binaries under different conditions.
@@
 
-\section{Common Envelope} \label{sec:ce}
+\section{Formation Through Common Envelope} \label{sec:ce}
 In \cite{yamaguchi_hi, yamaguchi_lo}, the author discussed the formation of WD+MS binaries through CE for both high-mass and low-mass systems, and it was found that wide WD+MS systems can be formed for certain energy budgets and certain initial separation in the MESA model.
@@
 To compare MESA model with COSMIC, we intend to compare how the energy budget utilized by \cite{yamaguchi_hi, yamaguchi_lo} matches with the CE model in COSMIC.
-We approach this by calculating the effective $\lambda$ value for each of the energy budget used in \cite{yamaguchi_hi, yamaguchi_lo}.
+We approach this by calculating the effective $\lambda$ value for each of the common envelope energy budget used in \cite{yamaguchi_hi, yamaguchi_lo}.
 
@@
 
-In \cite{yamaguchi_hi}, the lower limit of WD mass $M_{\mathrm{MD, min}}$ in the observational sample is calculated to be $1.244\Msun \sim 1.418\Msun$.
+According to \cite{yamaguchi_hi}, the lower limit of WD mass $M_{\mathrm{MD, min}}$ in the observational sample is calculated to be $1.244\Msun \sim 1.418\Msun$.
 The corresponding mass of MS progenitor is expected to be in the range of $6 \sim 9 \Msun$, and the companion mass has median around $1\Msun$.
@@
 We run COSMIC binary evolution simulation of a $7\Msun + 1\Msun$ system and investigate how the final separation $a_f$ depends on the initial separation $a_i$ and the CE parameter $\lambda$.
-We vary the initial separation in range $2 \sim 8 \au$ with $400$ steps.
-We vary the CE flag \verb|lambdaf| in range $0 \sim -100$, which corresponds to $\lambda = 0 \sim 100$, with $2000$ steps.
-Also, we run these parameters with four different CE efficiencies — $\alphace = 1$, $\alphace = 0.9$, $\alphace = 0.6$, and $\alphace = 0.3$, the same as documented in \cite{yamaguchi_hi}.
-In total, for $4$ different CE efficiency $\alphace$, we run a grid of $80000$ binary star systems with different initial separation $a_i$ and CE parameter $\lambda$.
+The initial separation is a linear grid in range $2 \sim 8 \au$ with $400$ points.
+The CE flag \verb|lambdaf| is a linear grid in range $0 \sim -100$ with $2000$ steps, which corresponds to $0 \sim 100$ for CE parameter $\lambda$.
+Also, we run these parameters with four different CE efficiencies — $\alphace = 1$, $\alphace = 0.9$, $\alphace = 0.6$, and $\alphace = 0.3$, the same as recorded in \cite{yamaguchi_hi}.
+In total, for each of the $4$ different CE efficiency $\alphace$, we run a grid of $80000$ binary star systems with different initial separation $a_i$ and CE parameter $\lambda$.
 
-After simulation is finished, we select the time-step when the system finished CE for the first time (\verb|evol_type = 8|).
+After simulation is finished, we select the time-step when the system finished CE for the first time (\verb|evol_type=8|).
 At this moment, if the system is a WD+MS system, we record their separation as the final separation $a_f$.
 Otherwise, we set $a_f = 0$ and consider this initial condition unable to produce a WD+MS system.
-
 We create four heat maps for the four different CE efficiencies.
-In Figure \ref{res_hi}, we plot the final separation $a_f$ of the WD+MS systems against initial separation $a_i$ and the CE flag \verb|lambdaf|$= - \lambda$.
+In Figure \ref{res_hi}, we plot the final separation $a_f$ of the WD+MS systems against initial separation $a_i$ and the CE flag $- \lambda$.
 
@@
   \centering
-  \includegraphics[width=0.8\linewidth]{fig/7+1results.png}
-  \caption{ Heat map of final separation $a_f$ against initial separation $a_i$ and CE flag lambdaf$=-\lambda$.
+  \includegraphics[width=\linewidth]{fig/7+1results.png}
+  \caption{ Heat map of final separation $a_f$ against initial separation $a_i$ and CE flag $-\lambda$.
 Different panels represent four different CE efficiencies $\alphace = 1$, $0.9$, $0.6$, and $0.3$.
@@
 There are several interesting points of the results.
-First of all, as $\alphace$ decreases, the final separation decreases in general.
+First of all, comparing the four panels, we found that as $\alphace$ decreases, the final separation decreases in general.
 This can be explained by Equation \ref{ebind-sep}.
@@
 That is, during the evolution process, two CE processes take place, between which there is a period of time.
-By checking the \verb|bpp| array, which documents the evolution processes of the binary star systems, we found that (***TO-DO***) fraction of all our simulated binaries experience two CE processes before reaching WD+MS.
-When the primary loses too much mass during the first CE, the CE process ends.
+A typical change of \verb|evol_type| through the evolution is \textbf{\dots-3-7-8-4-\dots-3-7-8-4-\dots}, where WD+MS is reached during the second mass transfer process.
+Another possible change of \verb|evol_type| through the volution is \textbf{\dots-3-7-8-7-8-4-\dots}, which will be mentioned in the next paragraph.
+Statistically, we found that \textbf{(***TO-DO***)} fraction of all our $320000$ simulated binaries experience two CE processes before reaching WD+MS.
+To explain this qualitatively, we notice that when the primary loses too much mass during the first CE, the CE process ends, as the radius of the primary becomes smaller than the Roche Lobe.
 After the primary fills its Roche lobe and has a large enough envelope again, the CE process resumes and causes two CE processes in total.
 
-From Figure \ref{res_hi}, we also notice that the final separation $a_f$ jumps obviously at about $2.9 \au$, and $7.1 \au$.
+From Figure \ref{res_hi}, we also notice that final separation $a_f$ jumps obviously at about $2.9 \au$, and $7.1 \au$.
 After investigating the \verb|bpp| array of these simulations, we find out the jump at about $2.9 \au$ and $7.1 \au$ is likely due to the difference of \verb|kstar_1| at the start of RLOF.
-For $a_i < 2.9 \au$, RLOF starts when the primary is on the giant branch (GB) with \verb|kstar_1 3|.
-For $a_i = 2.9 \au$, RLOF starts when the primary is experiencing core Helium burning (CHeB) with \verb|kstar_1 = 4|.
-For $2.9 \au < a_i < 7.1 \au$, RLOF starts when the primary is on E-AGB with \verb|kstar_1 = 5|.
-For $a_i > 7.1 \au$, RLOF starts when the primary is on TP-AGB with \verb|kstar_1 = 6|.
+For $a_i < 2.9 \au$, RLOF starts when \verb|kstar_1=3|.
+For $a_i = 2.9 \au$, RLOF starts when \verb|kstar_1=4|.
+For $2.9 \au < a_i < 7.1 \au$, RLOF starts when \verb|kstar_1=5|.
+For $a_i > 7.1 \au$, RLOF starts when \verb|kstar_1=6|.
 For different star type, the structure and mass of the envelope is different, leading to different binding energy $\Ebind$ and thus different final separation $a_f$.
@@
   \centering
-  \includegraphics[width = 0.5\linewidth]{fig/jump-zoom.png}
+  \includegraphics[width = 0.6\linewidth]{fig/jump-zoom.png}
   \caption{Dependence of final separation $a_f$ on initial separation $a_i$ for $\alphace = 0.3$.
@@
 
-Inspecting more closely, we find that the final separation $a_f$ jumps at about $5.2 \sim 5.4 \au$.
+Inspecting more closely, we found that the final separation $a_f$ jumps at about $5.2 \sim 5.4 \au$.
 This is especially obvious with $\alphace = 0.3$.
@@
 To explain this, we again go to the \verb|bpp| array of these systems.
-We notice that the first jump is probably because the evolution process changes from \textbf{3-7-8-4-3-7-8-4} to \textbf{3-7-8-7-8-4}.
+The first jump is because the evolution process changes from \textbf{3-7-8-4-3-7-8-4} to \textbf{3-7-8-7-8-4}.
 This means that for relatively larger initial separation $a_i$, the primary fills its Roche lobe all the time, leading to more mass loss during a shorter timescale and thus larger final separation.
-Also, the second jump is probably because \verb|kstar_1| changes from \verb|8| (Naked Helium Star Hertzsprung Gap) to \verb|9| (Naked Helium Star Giant Branch) at the start of the second common envelope.
+For the second jump, it is because \verb|kstar_1| changes from \verb|8| to \verb|9| at the start of the second common envelope.
 This changes the binding energy of the common envelope and thus changes the final separation.
@@
   \centering
-  \includegraphics[width=0.9\linewidth]{fig/7+1fit.png}
+  \includegraphics[width=\linewidth]{fig/7+1fit.png}
   \caption{Effective $\lambda$ fitting results of a $7\Msun + 1\Msun$ system for common envelope structure in MESA.
+Different lines represent different $\alphace$ and different common envelope energy budget, as shown in the legend.
 Left panel: MESA results in \cite{yamaguchi_hi} of how final separation $a_f$ depends on initial separation $a_i$.
@@
   \centering
-  \includegraphics[width=0.9\linewidth]{fig/7+1cmp.png}
+  \includegraphics[width=\linewidth]{fig/7+1cmp.png}
   \caption{Same as Figure \ref{fit_hi}, but with default $\lambda$ in COSMIC model included in the right panel.}
@@
 Notice that the default $\lambda$ value is at order $\sim 1$, far smaller than the effective $\lambda$ needed to recreate results in \cite{yamaguchi_hi}.
+Hence, we conclude that in COSMIC, it is not practical to form wide high mass WD+MS system through common envelope processes, since the required $\lambda$ is too large.
 
@@
   \centering
-  \includegraphics[width=0.9\linewidth]{fig/7+1ebcmp.png}
-  \caption{Same as Figure \ref{fit_cmp_hi}, but with a different energy budget — $\Ebind = \Egrav + \alphath \Eth + \alpharec \Erec$.}
+  \includegraphics[width=\linewidth]{fig/7+1ebcmp.png}
+  \caption{Same as Figure \ref{fit_cmp_hi}, but with a different common envelope energy budget.
+For results shown in these figures, $\Ebind = \Egrav + \alphath \Eth + \alpharec \Erec$, where the coefficients $\alphath$ and $\alpharec$ for different line is shown in the legend.}
   \label{fit_cmp_eb_hi}
@@
 
-In Figure \ref{fit_cmp_eb_hi}, we present the same results for another energy budget of the CE process — $\Ebind = \Egrav + \alphath \Eth + \alpharec \Erec$.
+In Figure \ref{fit_cmp_eb_hi}, we present the same results for another energy budget of the CE process.
+Rather than including both thermal and recombination energy into the binding energy, we use two coefficients $\alphath$ and $\alpharec$ to consider their contribution to $\Ebind$ separately.
+That is,
+\[
+  \Ebind = \Egrav + \alphath \Eth + \alpharec \Erec.
+\]
+We notice that the effective $\lambda$ now gets with in the order of $\sim 10$.
+However, none of these initial separation and $\lambda$ results in wide WD+MS system with final separation $a_f > 0.15 \au$.
+Therefore, we conclude that with the new energy budget, it is not practical to form wide WD+MS binaries either.
 
@@
 
-\section{Stable Mass Transfer} \label{sec:stable}
+\section{Formation Through Stable Mass Transfer} \label{sec:stable}
 
-\section{Population} \label{sec:population}
+\section{Population Synthesis Results} \label{sec:population}
 
 \bibliography{reference.bib}
-\bibliographystyle{apalike}
+\bibliographystyle{alpha}
 

### 2024-07-08 summaries/paper-draft-plain.tex
finished low mass results section

@@
 
-In \cite{yamaguchi_hi, yamaguchi_lo}, to explore the dependency of the final separation $a_f$ in the post-CE binary on the initial separation $a_i$ in the pre-MT binary, the author first run a stellar evolution model of the primary star using MESA.
+In \cite{yamaguchi_hi, yamaguchi_lo}, to explore the dependency of the final separation $a_f$ in the post-CE binary on the initial separation $a_i$ in the pre-MT binary, the author first ran a stellar evolution model of the primary star using MESA.
 Its evolution up to AGB phase is followed.
@@
 This $\lambda$ is represented in COSMIC as the \verb|lambdaf| flag in \verb|BSEDict|.
-To compare MESA model with COSMIC, we intend to compare how the energy budget utilized by \cite{yamaguchi_hi, yamaguchi_lo} matches with the CE model in COSMIC.
+To compare MESA model with COSMIC, we intend to compare how the energy budget utilized in past papers matches with the CE model in COSMIC.
 We approach this by calculating the effective $\lambda$ value for each of the common envelope energy budget used in \cite{yamaguchi_hi, yamaguchi_lo}.
@@
 
-We run COSMIC binary evolution simulation of a $7\Msun + 1\Msun$ system and investigate how the final separation $a_f$ depends on the initial separation $a_i$ and the CE parameter $\lambda$.
+We ran COSMIC binary evolution simulation of a $7\Msun + 1\Msun$ system and investigate how the final separation $a_f$ depends on the initial separation $a_i$ and the CE parameter $\lambda$.
 The initial separation is a linear grid in range $2 \sim 8 \au$ with $400$ points.
 The CE flag \verb|lambdaf| is a linear grid in range $0 \sim -100$ with $2000$ steps, which corresponds to $0 \sim 100$ for CE parameter $\lambda$.
-Also, we run these parameters with four different CE efficiencies — $\alphace = 1$, $\alphace = 0.9$, $\alphace = 0.6$, and $\alphace = 0.3$, the same as recorded in \cite{yamaguchi_hi}.
-In total, for each of the $4$ different CE efficiency $\alphace$, we run a grid of $80000$ binary star systems with different initial separation $a_i$ and CE parameter $\lambda$.
+Also, we ran these parameters with four different CE efficiencies — $\alphace = 1$, $\alphace = 0.9$, $\alphace = 0.6$, and $\alphace = 0.3$, the same as recorded in \cite{yamaguchi_hi}.
+In total, for each of the $4$ different CE efficiency $\alphace$, we ran a grid of $80000$ binary star systems with different initial separation $a_i$ and CE parameter $\lambda$.
 
-After simulation is finished, we select the time-step when the system finished CE for the first time (\verb|evol_type=8|).
+After simulation is finished, we select the time-step when the system finished CE for the first time (\verb|evol_type = 8|).
 At this moment, if the system is a WD+MS system, we record their separation as the final separation $a_f$.
@@
   \includegraphics[width=\linewidth]{fig/7+1results.png}
-  \caption{ Heat map of final separation $a_f$ against initial separation $a_i$ and CE flag $-\lambda$.
+  \caption{Heat map of final separation $a_f$ against initial separation $a_i$ and CE flag $-\lambda$.
 Different panels represent four different CE efficiencies $\alphace = 1$, $0.9$, $0.6$, and $0.3$.
@@
 After investigating the \verb|bpp| array of these simulations, we find out the jump at about $2.9 \au$ and $7.1 \au$ is likely due to the difference of \verb|kstar_1| at the start of RLOF.
-For $a_i < 2.9 \au$, RLOF starts when \verb|kstar_1=3|.
-For $a_i = 2.9 \au$, RLOF starts when \verb|kstar_1=4|.
-For $2.9 \au < a_i < 7.1 \au$, RLOF starts when \verb|kstar_1=5|.
-For $a_i > 7.1 \au$, RLOF starts when \verb|kstar_1=6|.
+For $a_i < 2.9 \au$, RLOF starts when \verb|kstar_1 = 3|.
+For $a_i = 2.9 \au$, RLOF starts when \verb|kstar_1 = 4|.
+For $2.9 \au < a_i < 7.1 \au$, RLOF starts when \verb|kstar_1 = 5|.
+For $a_i > 7.1 \au$, RLOF starts when \verb|kstar_1 = 6|.
 For different star type, the structure and mass of the envelope is different, leading to different binding energy $\Ebind$ and thus different final separation $a_f$.
@@
 
-In conclusion, with a large \verb|lambdaf|, it is possible to create wide post-mass transfer WD+MS systems with final separation $a_f > 0.15 \au$, which is the minimal separation of our observed objects.
+In conclusion, with a large $\lambda$, it is possible to create wide post-mass transfer WD+MS systems with final separation $a_f > 0.15 \au$, which is the minimal separation of our observed objects.
 In next section, we will compare these results with default settings in COSMIC and with results in \cite{yamaguchi_hi}, in order to investigate the possibility of forming wide WD+MS systems through CE.
@@
 After we have obtained the simulation results for different initial conditions, now we can calculate the effective $\lambda$ that matches with results in \cite{yamaguchi_hi}.
-For each fixed initial separation $a_i$, we loop through the COSMIC results with the same initial separation, and find the \verb|lambdaf| value that results in a final separation closest to MESA model results.
-We record this \verb|lambdaf| as the effective $\lambda$ that matches the binding energy formalism at initial separation $a_i$.
+For each fixed initial separation $a_i$, we loop through the COSMIC results with the same initial separation, and find the $\lambda$ value that results in a final separation closest to MESA model results.
+We record this $\lambda$ as the effective $\lambda$ that matches the binding energy formalism at initial separation $a_i$.
 
@@
 Different lines represent different $\alphace$ and different common envelope energy budget, as shown in the legend.
-Left panel: MESA results in \cite{yamaguchi_hi} of how final separation $a_f$ depends on initial separation $a_i$.
+Left panel: recreation of MESA results in \cite{yamaguchi_hi} of how final separation $a_f$ depends on initial separation $a_i$.
 Central panel: COSMIC results of how final separation $a_f$ depends on initial separation $a_i$, produced by the effective $\lambda$ calculated.
@@
 From the left panel, we can see that it is possible to create WD+MS binaries with final separation $>0.15 \au$ in MESA as long as internal energy $\Eint$ is included.
-However, in the right panel we show that this correspond to very large \verb|lambdaf| in COSMIC ($\lambda > \sim 20$).
+However, in the right panel we show that this correspond to very large $\lambda$ in COSMIC ($\lambda > \sim 20$).
 
@@
   \includegraphics[width=\linewidth]{fig/7+1cmp.png}
-  \caption{Same as Figure \ref{fit_hi}, but with default $\lambda$ in COSMIC model included in the right panel.}
+  \caption{Same as Figure \ref{fit_hi}, but with default $\lambda$ in COSMIC model included in the right panel with the purple line.}
   \label{fit_cmp_hi}
@@
   \caption{Same as Figure \ref{fit_cmp_hi}, but with a different common envelope energy budget.
-For results shown in these figures, $\Ebind = \Egrav + \alphath \Eth + \alpharec \Erec$, where the coefficients $\alphath$ and $\alpharec$ for different line is shown in the legend.}
+For results shown in these figures, $\Ebind = \Egrav + \alphath \Eth + \alpharec \Erec$, where the coefficients $\alphath$ and $\alpharec$ for different line is shown in the legend.
+The red line in the right panel indicates default $\lambda$ in COSMIC.}
   \label{fit_cmp_eb_hi}
@@
 
-\subsection{Low Mass Systems $(1.5 \Msun + 0.85 \Msun)$} \label{subsec:low}
-In \cite{yamaguchi_lo}, the WD mass in the Gaia sample ranges from $0.5\Msun$ to $0.8\Msun$, which corresponds to progenitor mass of $1 \Msun$ to $3 \Msun$.
+\subsection{Low Mass Systems ($1.5 \Msun + 0.85 \Msun$)} \label{subsec:low}
+
+\subsubsection{Methods and results}
+According to \cite{yamaguchi_lo}, the WD mass in the Gaia sample ranges from $0.5\Msun$ to $0.8\Msun$, which corresponds to progenitor mass of $1 \Msun$ to $3 \Msun$.
 The companions mass has a median of $0.85 \Msun$.
@@
 Hence, we consider a $1.5\Msun + 0.85\Msun$ evolution model in COSMIC and compare the results with those in \cite{yamaguchi_hi}.
+For the common envelope, we include all thermal energy $\Eth$ into the binding energy $\Ebind$, while the recombination energy $\Erec$ is included in part with a parameter $\alpharec$.
+That is,
+\[
+  \Ebind = \Egrav + \Eth + \alpharec \Erec.
+\]
+
+We take the same approach as high mass systems to investigate the effective value of $\lambda$ in COSMIC that matches with the results in MESA.
+The initial separation is a linear grid in range $0.5 \sim 6 \au$ with $400$ points.
+The CE flag \verb|lambdaf| is a linear grid in range $0 \sim -100$ with $2000$ steps, which corresponds to $0 \sim 100$ for CE parameter $\lambda$.
+Also, we ran these parameters with four different CE efficiencies — $\alphace = 1$, $\alphace = 0.9$, $\alphace = 0.6$, and $\alphace = 0.3$, the same as recorded in \cite{yamaguchi_lo}.
+In total, for each of the $4$ different CE efficiency $\alphace$, we ran a grid of $80000$ binary star systems with different initial separation $a_i$ and CE parameter $\lambda$.
+
+After simulation is finished, we again select the time-step when the system finished CE for the first time, and record the final separation if the system is a desried WD+MS system.
+The result is presented in Figure \ref{res_lo}.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{fig/1.5+0.8results.png}
+  \caption{Heat map of final separation $a_f$ against initial separation $a_i$ and CE flag $-\lambda$.
+Different panels represent four different CE efficiencies $\alphace = 1$, $0.9$, $0.6$, and $0.3$.
+The white and yellow contour corresponds to $a_f = 0.01 \au$ and $a_f = 0.15 \au$ respectively.
+WD+MS binaries is not possible for separation smaller than $0.01 \au$, and $0.15 \au$ is the minimum separation of wide WD+MS binaries in observational results.
+Outside the white contour no desired system forms.}
+  \label{res_lo}
+\end{figure}
 
@@
 \bibliography{reference.bib}
-\bibliographystyle{alpha}
+\bibliographystyle{apalike}
 

### 2024-07-09 summaries/paper-draft-plain.tex
continued summary and updated graph

@@
 In section \ref{sec:population}, we will research on how different COSMIC model affect simulation results of a sample population of binary star systems, and how the formation of post WD+MS binaries depends on various parameters.
+
+Throughout the paper, the stellar type and evolution type follows the BSE conventions.
+\verb|kstar| values specify the evolutionary state of the star, and \verb|evol_type| values specify the evolutionary changes of the binary systems:
+
+\begin{tabular}{cl|cl}
+  \verb|kstar| & evolutionary state & \verb|evol_type| & evolutionary change\\
+  \hline
+  0 & MS $<0.7 \Msun$ & 1 & initial state \\
+  1 & MS $>0.7 \Msun$ & 2 & kstar change \\
+  2 & Hertzprung Gap & 3 & begin Roche lobe overflow\\
+  3 & First Giant Branch & 4 & end Roche lobe overflow\\
+  4 & Core Helium Burning & 5 & contact \\
+  5 & Early AGB & 6 & coalescence \\
+  6 & TP-AGB & 7 & begin common envelope \\
+  7 & Naked Helium Star MS & 8 & end common envelope \\
+  8 & Naked Helium Star Hertzprung Gap & 9 & no remnant leftover \\
+  9 & Naked Helium Star Giant Branch & 10 & max evolution time \\
+  10 & Helium White Dwarf & 11 & binary disruption \\
+  11 & Carbon/Oxygen White Dwarf & 12 & begin symbiotic phase \\
+  12 & Oxygen/Neon White Dwarf & 13 & end symbiotic phase \\
+  13 & Neutron Star & 14 & blue straggler \\
+  14 & Black Hole & 15 & supernova of primary \\
+  15 & Massless Remnant & 16 & supernova of secondary
+\end{tabular}
 
@@
 Statistically, we found that \textbf{(***TO-DO***)} fraction of all our $320000$ simulated binaries experience two CE processes before reaching WD+MS.
-To explain this qualitatively, we notice that when the primary loses too much mass during the first CE, the CE process ends, as the radius of the primary becomes smaller than the Roche Lobe.
+To explain this qualitatively, we notice that when the primary loses too much mass during the first CE, the CE process ends, as the radius of the primary becomes smaller than the Roche lobe.
 After the primary fills its Roche lobe and has a large enough envelope again, the CE process resumes and causes two CE processes in total.
@@
 This is because at $6 \sim 7 \au$, we cannot reach such low final separation in COSMIC while forming a WD+MS binary at the same time.
-From the left panel, we can see that it is possible to create WD+MS binaries with final separation $>0.15 \au$ in MESA as long as internal energy $\Eint$ is included.
+From the left panel, we can see that it is possible to create WD+MS binaries with $a_f > 0.15 \au$ in MESA as long as internal energy $\Eint$ is included, with initial separation $a_i > 3.5 \au$.
 However, in the right panel we show that this correspond to very large $\lambda$ in COSMIC ($\lambda > \sim 20$).
@@
 
+To compare the results with \cite{yamaguchi_lo}, we use the same approach as that for high mass systems.
+In Figure \ref{fit_cmp_lo_whole}, we compare COSMIC with MESA model that includes binding energy of the entire envelope.
+In Figure \ref{fit_cmp_lo_pt}, we compare COSMIC with MESA model that includes only binding energy of the outer envelope.
+Notice that we do not have fitting results for $a_i > 4 \au$.
+This is because no WD+MS binaries form in COSMIC for such initial separation range.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{fig/1.5+0.8cmp_whole.png}
+  \caption{Effective $\lambda$ fitting results of a $1.5\Msun + 0.85\Msun$ system for common envelope structure in MESA.
+Binding energy of the \emph{entire envelope} is considered.
+Left Panel: MESA results in \cite{yamaguchi_lo} of how final separation $a_f$ depends on initial separation $a_i$.
+Central Panel: COSMIC results of how final separation $a_f$ depends on initial separation $a_i$, produced by the effective $\lambda$.
+Right Panel: effective $\lambda$ in COSMIC that corresponds to the MESA results for each initial separation $a_i$.
+The purple line represents default $\lambda$ calculated using \cite{claeys2014theoretical}.
+The red dashed line and the blue dashed line in the left and central panel represents $0.15 \au$ and $0.7 \au$ respectively.
+$0.15 \au$ is the minimal separation for observed wide post-CE WD+MS binaries, while $0.7 \au$ is the typical separation for wide post-CE WD+MS binaries.}
+  \label{fit_cmp_lo_whole}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{fig/1.5+0.8cmp_pt.png}
+  \caption{Same as Figure \ref{fit_cmp_lo_pt} but only binding enregy of the \emph{outer envelope} is considered.}
+  \label{fit_cmp_lo_pt}
+\end{figure}
+
+From the left panel of Figure \ref{fit_cmp_lo_whole}, we can see that wide WD+MS systems can form in MESA model when $a_i > 2.5 \au$.
+This is reproduced in COSMIC with $\lambda \sim 1$.Hence, wide WD+MS can form for $a_i > 2.5 \au$ with parameter $\lambda$ about the same as default.
+Recall that for high mass system, wide WD+MS systems cannot form without $\lambda \sim 10^2$.
+This is a significant difference between high mass systems and low mass systems.
+
 \section{Formation Through Stable Mass Transfer} \label{sec:stable}
+
+After exploring the formation of WD+MS binaries through common envelope, we investigate the possibility of forming WD+MS binaries through stable mass transfer.
+In \cite{yamaguchi_lo}, the author plot the $\MWD - \Porb$ relation of observed wide WD+MS bianries, and argues that it is not possible to form wide WD+MS binaries through stable mass.
+The plot is recreated in Figure \ref{theory-observed}, and from the figure, we can see that the observed wide post-mass transfer WD+MS bianries do not match the theoretical $\MWD - \Porb$ relation for stable mass transfer calculated in \cite{rappaport1995relation}.
+Also, the author aruges that stable mass transfer requires donors on SGB at the onset of mass transfer.
+However, this will not lead to WDs with masses observed \cite{yamaguchi_lo}.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.8\linewidth]{fig/theory-observed.png}
+  \caption{$M_{\mathrm{WD}} - P_{\mathrm{orb}}$ relation of observed wide post-mass transfer binaries.
+White dwarf mass $\MWD$ correspond to The colors of the points represent the masses of the luminous companion, $\Mstar$.
+On the right axis, we show the minimum orbital separation $a_{\mathrm{peri}}$, corresponding to the period assuming companion mass $\Mstar = 1 \Msun$ and white dwarf mass $\MWD = 0.6 \Msun$.
+The dashed line and the grey region is the track along which binaries undergoing stable MT are expected to evolve according to \cite{rappaport1995relation}}.
+  \label{theory-observed}
+\end{figure}
+
+As in the common envelope section, we investigate how the formation of WD+MS binaries depends on various parameters in high mass systems using $7\Msun + 1 \Msun$ binaries, and in low mass systems using $1.5\Msun + 0.85\Msun$ binaries.
 

### 2024-07-10 summaries/paper-draft-plain.tex
change paper draft stable mass section

@@
 
-\begin{tabular}{cl|cl}
-  \verb|kstar| & evolutionary state & \verb|evol_type| & evolutionary change\\
+\begin{tabular}{cl | cl}
+  \verb|kstar| & \textbf{evolutionary state} & \verb|evol_type| & \textbf{evolutionary change} \\
   \hline
@@
 
+It is worth noting that most of these systems experience two mass transfer phases in COSMIC before reaching WD+MS.
+That is, during the evolution process, two CE processes take place, between which there is a period of time.
+A typical change of \verb|evol_type| through the evolution is \textbf{\dots-3-7-8-4-\dots-3-7-8-4-\dots}, where WD+MS is reached during the second mass transfer process.
+Another possible change of \verb|evol_type| through the volution is \textbf{\dots-3-7-8-7-8-4-\dots}, which will be mentioned in the next paragraph.
+Statistically, we found that \textbf{(***TO-DO***)} fraction of all our $320000$ simulated binaries experience two CE processes before reaching WD+MS.
+To explain this qualitatively, we notice that when the primary loses too much mass during the first CE, the CE process ends, as the radius of the primary becomes smaller than the Roche lobe.
+After the primary fills its Roche lobe and has a large enough envelope again, the CE process resumes and causes two CE processes in total.
+
 \begin{figure}
@@
 Qualitatively, note that more energy from the orbit is needed to eject the envelope for low $\alphace$, so the resulting $a_f$ is expected to be smaller.
-
-It is worth noting that most of these systems experience two mass transfer phases in COSMIC before reaching WD+MS.
-That is, during the evolution process, two CE processes take place, between which there is a period of time.
-A typical change of \verb|evol_type| through the evolution is \textbf{\dots-3-7-8-4-\dots-3-7-8-4-\dots}, where WD+MS is reached during the second mass transfer process.
-Another possible change of \verb|evol_type| through the volution is \textbf{\dots-3-7-8-7-8-4-\dots}, which will be mentioned in the next paragraph.
-Statistically, we found that \textbf{(***TO-DO***)} fraction of all our $320000$ simulated binaries experience two CE processes before reaching WD+MS.
-To explain this qualitatively, we notice that when the primary loses too much mass during the first CE, the CE process ends, as the radius of the primary becomes smaller than the Roche lobe.
-After the primary fills its Roche lobe and has a large enough envelope again, the CE process resumes and causes two CE processes in total.
 
@@
 
-\subsubsection{Methods and results}
+\subsubsection{Methods and Results}
 According to \cite{yamaguchi_lo}, the WD mass in the Gaia sample ranges from $0.5\Msun$ to $0.8\Msun$, which corresponds to progenitor mass of $1 \Msun$ to $3 \Msun$.
@@
 However, this will not lead to WDs with masses observed \cite{yamaguchi_lo}.
+We will verfiy this idea in this section using $7\Msun + 1 \Msun$ binaries.
 
@@
 
-As in the common envelope section, we investigate how the formation of WD+MS binaries depends on various parameters in high mass systems using $7\Msun + 1 \Msun$ binaries, and in low mass systems using $1.5\Msun + 0.85\Msun$ binaries.
+\subsection{Methods and Results}
+First of all, to ensure stable mass transfer in COSMIC, we modify every entry of \verb|qcrit_array| in \verb|BSEDict| from the default $0.0$ to $20.0$.
+This changes the critical mass ratio for onset of unstable mass transfer to $20.0$ Based on our initial mass ratio $7:1$, this will make the actual mass raio of our system below this critical throughout the evolutionary process.
+
+Secondly, a parameter on which stable mass transfer process depends is the accrection efficiency $\beta$, which in COSMIC is represented by \verb|acc_lim| in \verb|BSEDict|.
+Accretion efficiency is the fraction $\beta$ of all transfered materials that is accreted by the accretor.
+That is, $\dot{M_a} = -\beta\dot{M_d}$ where $M_d$ is the mass of donor and $M_a$ is the mass of accretor.
+
+In our simulation, we investigate how accretion efficiency $\beta$ and initial separation $a_i$ affects the final separation $a_f$.
+We run a grid of three different $\beta$ — $\beta = 0.0$, $0.5$ and $1.0$, and a linear grid of $a_i$ from $2 \au$ to $8 \au$ with $400$ steps.
+In total, we simulate $1200$ binaries with different accretion limit and initial separation.
+For each system, after the simulation is finished, we select the first timestep, if any, that shows a WD+MS binary.
+We record the separation at this moment as the final separation $a_f$, and we compare them with the observed minimum separation $0.15 \au$ for post-mass transfer WD binaries.
+
+The results are presented in Figure \ref{stable_hi}.
+Notice the gap for $a_i$ ranges in $3 \sim 4.5 \au$.
+This means no WD+MS forms for this initial separation.
+In fact, for these initial separations and accretion efficiencies, two star merges.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.75\linewidth]{fig/stable_hi_log.png}
+  \caption{COSMIC results of WD+MS binaries after stable mass transfer for $7\Msun + 1\Msun$ systems, with accretion efficiency $\beta = 0$, $\beta = 0.5$, and $\beta = 1$ respectively.
+Left panel: dependence of final separation $a_f$ of the WD+MS binary on initial separation $a_i$.
+Right panel: The corresponding companion mass in solar units in the WD+MS binary.
+The gap in both figure represents that no WD+MS system forms for that initial separation.}
+  \label{stable_hi}
+\end{figure}
+
+Notice that the majority of these systems have final separation $a_f > 0.15 \au$.
+However, this is \textbf{not the correct approach}, since the companion mass in our observed wide WD+MS binaries has median mass $\sim 1\Msun$.
+However, in the simulation results shown in Figure \ref{stable_hi}, the final companion mass is about $4 \Msun$ and $6.5 \Msun$ for $\beta = 0.5$ and $\beta = 1.0$ respectively.
+Hence, we need to adjust the  initial companion mass $M_{\star, i}$ and the accretion efficiency \verb|acc_lim|, in order to make the final companion mass $M_{\star, f}$ close to $1\Msun$.
+Therefore, we lower the initial companion mass $M_{\star, i}$ and adjust the accretion efficiency $\beta$ and re-run the simulation.
+The results for $7\Msun + 0.7\Msun$ with $\beta = 0.1$, and $7\Msun + 0.5\Msun$ with $\beta = 0.2$ is shown in Figure \ref{stable_hi_ex}.
+After adjusting the initial mass of companion and the accretion efficiency, now the final mass of the companion is about $1\Msun$ and we get systems similar to our observed sample.
+
+\begin{figure}
+  \centering
+  \includegraphics[width = 0.75\linewidth]{fig/stable7+1ex.png}
+  \caption{COSMIC results of WD+MS binaries after stable mass transfer for two systems — $7\Msun + 0.7\Msun$ with accretion efficiency $\beta = 0.1$, and $7\Msun + 0.5\Msun$ with accretion efficiency $\beta = 0.2$.
+The left and right panel presents the same content as in Figure \ref{stable_hi}.}
+  \label{stable_hi_ex}
+\end{figure}
 

### 2024-07-10 summaries/paper-draft-plain.tex
edited stable MT section and start population section

@@
 \usepackage{listings}
+\usepackage{amsmath}
 \usepackage[labelfont=bf]{caption}
@@
 
-\begin{tabular}{cl | cl}
+\begin{tabular}{|cl|cl|}
+  \hline
   \verb|kstar| & \textbf{evolutionary state} & \verb|evol_type| & \textbf{evolutionary change} \\
@@
   14 & Black Hole & 15 & supernova of primary \\
-  15 & Massless Remnant & 16 & supernova of secondary
+  15 & Massless Remnant & 16 & supernova of secondary \\
+  \hline
 \end{tabular}
@@
 
-It is worth noting that most of these systems experience two mass transfer phases in COSMIC before reaching WD+MS.
-That is, during the evolution process, two CE processes take place, between which there is a period of time.
-A typical change of \verb|evol_type| through the evolution is \textbf{\dots-3-7-8-4-\dots-3-7-8-4-\dots}, where WD+MS is reached during the second mass transfer process.
-Another possible change of \verb|evol_type| through the volution is \textbf{\dots-3-7-8-7-8-4-\dots}, which will be mentioned in the next paragraph.
-Statistically, we found that \textbf{(***TO-DO***)} fraction of all our $320000$ simulated binaries experience two CE processes before reaching WD+MS.
-To explain this qualitatively, we notice that when the primary loses too much mass during the first CE, the CE process ends, as the radius of the primary becomes smaller than the Roche lobe.
-After the primary fills its Roche lobe and has a large enough envelope again, the CE process resumes and causes two CE processes in total.
-
 \begin{figure}
@@
 \subsubsection{Discussions}
+
+A typical change of \verb|evol_type| through the evolution is $\boxed{\text{\textbf{*-3-7-8-4-*-3-7-8-4-*}}}$ (refer to Section \ref{sec:intro} for correspondence between \verb|evol_type| and evolutionary changes), where WD+MS is reached during the second mass transfer process.
+Another possible change of \verb|evol_type| through the volution is $\boxed{\text{\textbf{*-3-7-8-7-8-4-*}}}$.
+It is worth noting that most of these systems experience two mass transfer phases in COSMIC before reaching WD+MS.
+That is, during the evolution process, two CE processes take place, between which there is a period of time.
+Statistically, we found that \textbf{(***TO-DO***)} fraction of all our $320000$ simulated binaries experience two CE processes before reaching WD+MS.
+To explain this qualitatively, we notice that when the primary loses too much mass during the first CE, the CE process ends, as the radius of the primary becomes smaller than the Roche lobe.
+After the primary fills its Roche lobe and has a large enough envelope again, the CE process resumes and causes two CE processes in total.
 
@@
 To explain this, we again go to the \verb|bpp| array of these systems.
-The first jump is because the evolution process changes from \textbf{3-7-8-4-3-7-8-4} to \textbf{3-7-8-7-8-4}.
+The first jump is because the evolution process changes from $\boxed{\text{\textbf{3-7-8-4-3-7-8-4}}}$ to $\boxed{\text{\textbf{3-7-8-7-8-4}}}$.
 This means that for relatively larger initial separation $a_i$, the primary fills its Roche lobe all the time, leading to more mass loss during a shorter timescale and thus larger final separation.
@@
 However, in the simulation results shown in Figure \ref{stable_hi}, the final companion mass is about $4 \Msun$ and $6.5 \Msun$ for $\beta = 0.5$ and $\beta = 1.0$ respectively.
-Hence, we need to adjust the  initial companion mass $M_{\star, i}$ and the accretion efficiency \verb|acc_lim|, in order to make the final companion mass $M_{\star, f}$ close to $1\Msun$.
+Hence, we need to adjust the  initial companion mass $M_{\star, i}$ and the accretion efficiency $\beta$, in order to make the final companion mass $M_{\star, f}$ close to $1\Msun$.
 Therefore, we lower the initial companion mass $M_{\star, i}$ and adjust the accretion efficiency $\beta$ and re-run the simulation.
 The results for $7\Msun + 0.7\Msun$ with $\beta = 0.1$, and $7\Msun + 0.5\Msun$ with $\beta = 0.2$ is shown in Figure \ref{stable_hi_ex}.
-After adjusting the initial mass of companion and the accretion efficiency, now the final mass of the companion is about $1\Msun$ and we get systems similar to our observed sample.
+
+After adjusting the initial mass of companion and the accretion efficiency, now the final mass of the companion is about $1\Msun$ and we get systems similar to our observed samples. 
+Also, in Figure \ref{stable_hi_ex}, we note that when initial separation $a_i$ lies in the range of $2.5 \sim 4.5 \au$, the final separation incerases sharply.
+Also, it is only in this region that $a_f$ goes larger than the observed minimum $0.15 \au$ for wide WD binaries.
+Hence, in the next subsection, we will try to explain this sharp increases qualitatively.
 
@@
 
+\subsection{Discussions}
+In the previous section, we mentioned how $7\Msun + 1 \Msun$ systems with $\beta = 0$, $7 \Msun + 0.7\Msun$ systems with $\beta = 0.1$, and $7 \Msun + 0.5 \Msun$ systems with $\beta = 0.2$ produces WD+MS binaries with final separation $a_f > 0.15 \au$ and companion mass $\sim 1 \Msun$ in certain range of $a_i$.
+
+From figure \ref{stable_hi_ex}, we can see that the dependence of $a_f$ on $a_i$ can be roughly divided into four parts.
+The \textbf{pre-spike} part before the sharp increase with $a_i < 2.5 \au$, the \textbf{spike} part where the $a_f$ inrceases sharply with $2.5 \au < a_i < 3.5 \au$, the gap where the two stars merge with $3.5 \au < a_i < 4.5 \au$, and the part after that with $a_i > 4.5 \au$.
+
+Typical \textbf{pre-spike} and \textbf{spike} systems both follows the evolution path $\boxed{\text{\textbf{*-3-*-5-7-8-4-*-3-*-4-*}}}$ before reacing WD+MS.
+Note that the binaries come into contact (\verb|evol_type = 5|), which follows by a common enevelope phase.
+After inspecting the evolutionary process more closely, we found that the second stable mass transfer tends to widens the orbit significantly.
+\textbf{(***TO-DO***)}
+
 \section{Population Synthesis Results} \label{sec:population}
+After investigating how the final separation depends on various parameters in individual binary star systems, we now shift our focus onto the distribution of final orbital period in a population of binaries (equivalent to final separation since $\Porb^2 \sim a^3$).
+
+\subsection{Methods}
+First of all, we sample the binary population using the independent sampler in COSMIC.
+We set the primary mass model to \verb|korupa01|, eccentricity model to \verb|uniform|, and orbital period model, together with the binary fraction model to \verb|moe19|.
+We sample a total number of about $10000$ binaries with solar metallicity.
+The distribution of primary mass and the orbital period in the sampled population is shown in the Figure \ref{sample-distribution}.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{fig/sample-distribution.png}
+  \caption{The distribution of primary mass, companion mass, and orbital period in the sampled population using independent sampler.
+Left panel: Distribution of primary mass and companion mass in log scale.
+Right panel: Distribution of orbital period $\Porb$ in log scale.}
+  \label{sample-distribution}
+\end{figure}
 

### 2024-07-11 summaries/paper-draft-plain.tex
finish organizing current summary

@@
   1 & MS $>0.7 \Msun$ & 2 & kstar change \\
-  2 & Hertzprung Gap & 3 & begin Roche lobe overflow\\
+  2 & Hertzsprung Gap & 3 & begin Roche lobe overflow\\
   3 & First Giant Branch & 4 & end Roche lobe overflow\\
@@
   7 & Naked Helium Star MS & 8 & end common envelope \\
-  8 & Naked Helium Star Hertzprung Gap & 9 & no remnant leftover \\
+  8 & Naked Helium Star Hertzsprung Gap & 9 & no remnant leftover \\
   9 & Naked Helium Star Giant Branch & 10 & max evolution time \\
@@
 In our simulation, we investigate how accretion efficiency $\beta$ and initial separation $a_i$ affects the final separation $a_f$.
-We run a grid of three different $\beta$ — $\beta = 0.0$, $0.5$ and $1.0$, and a linear grid of $a_i$ from $2 \au$ to $8 \au$ with $400$ steps.
+We run a grid of three different accretion efficiency — $\beta = 0.0$, $0.5$ and $1.0$, and a linear grid of $a_i$ from $2 \au$ to $8 \au$ with $400$ steps.
 In total, we simulate $1200$ binaries with different accretion limit and initial separation.
@@
 Left panel: dependence of final separation $a_f$ of the WD+MS binary on initial separation $a_i$.
-Right panel: The corresponding companion mass in solar units in the WD+MS binary.
+Right panel: the corresponding companion mass in solar units in the WD+MS binary.
 The gap in both figure represents that no WD+MS system forms for that initial separation.}
@@
   \caption{The distribution of primary mass, companion mass, and orbital period in the sampled population using independent sampler.
-Left panel: Distribution of primary mass and companion mass in log scale.
-Right panel: Distribution of orbital period $\Porb$ in log scale.}
+Left panel: distribution of primary mass and companion mass in log scale.
+Right panel: distribution of orbital period $\Porb$ in log scale.}
   \label{sample-distribution}
@@
 
+To investigate how the final white dwarf mass and orbital period depends on \verb|lambdaf|, we simulation the same sampled population with five different $\lambda$ values, $\lambda = 5$, $15$, $25$, $35$, $45$.
+After that, we select the first time-step that satisfies \verb|evol_type = 8|, \verb|kstar_1| is WD, and \verb|kstar_2| is MS.
+We record the white dwarf mass $\MWD$ and the final orbital period $P_{\mathrm{orb}, f}$ from this moment.
+We also select the moment when the first Roche lobe overflow starts, and record the initial orbital separation $P_{\mathrm{orb}, i}$ and \verb|kstar_1| type at this moment.
+These variables will later help us study how the orbit widens, with respect to different parameter $\lambda$ and initial separation $a_i$.
+
+\subsection{Results}
+\begin{figure}
+  \centering
+  \includegraphics[width=0.6\linewidth]{fig/Mwd-P scatter.png}
+  \caption{Scatter plot of white dwarf mass $\MWD$ and final orbital period after common envelope in log scale $\log_{10}\Porb$.
+Different colors indicates results for different $\lambda$ values.
+For each different $\lambda$ value the same initial population is used.}
+  \label{Mwd-P-scatter}
+\end{figure}
+
+To show how the change in $\lambda$ affects the final orbital period, we plot final orbital period $\Porb$ against white dwarf mass $\MWD$ in Figure \ref{Mwd-P-scatter}.
+From Figure \ref{Mwd-P-scatter}, we note that the final orbital period is larger in general for larger $\lambda$.
+This can be clearly seen from the figure for $\MWD$ ranges in $0.2 \sim 0.7 \au$.
+This is also expected based on Equation \ref{ebind}, since for larger $\lambda$, less energy is required to unbind the common envelope, leading to more energy after mass transfer and a larger orbital period.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{fig/kstar-map-rearr.png}
+  \caption{Each panel: for this $\lambda$ value in the title, scatter plot of white dwarf mass $\MWD$ and final orbital period after common envelope in log scale $\log_{10}\Porb$.
+Different colors represent different kstar type of the primary at the start of mass transfer.}
+  \label{kstar-map}
+\end{figure}
+
+To also show how the final separation $\Porb$ depends on \verb|kstar_1| type at the start of first RLOF, we plot $\Porb$ against $\MWD$ for different $\lambda$ separately, and use different color to distinguish between different \verb|kstar_1| type at the start of mass transfer.
+The plot is shown in Figure \ref{kstar-map}.
+In general, if the star is at later stages of evolution (larger \verb|kstar|) when mass transfer starts, then the resulting white dwarf mass $\MWD$ and $\Porb$ are also both larger.
+Note also that few systems start the mass transfer at \verb|kstar_1 = 2| (Hertzsprung gap) and \verb|kstar_1 = 4| (Core Helium Burning).
+This is because stars crosses Hertzsprung in a short timescale, leaving little time for interaction.
+Also, stellar radius is small when going through Core Helium Burning, reducing the possibility of interactions that starts when \verb|kstar_1 = 4|.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.6\linewidth]{fig/dif-kstar map.png}
+  \caption{Scatter plot of difference in orbital period $\Delta P = P_{\mathrm{orb}}^{\lambda = 35} - P_{\mathrm{orb}}^{\lambda = 5}$ against the white dwarf mass $\MWD$.
+Different colors indicates different kstar type of the primary at the start of mass transfer.}
+  \label{dif-kstar map}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{fig/dif-kstar two plot.png}
+  \caption{Left panel: scatter plot of initial orbital period at the start of RLOF $P_{\mathrm{orb}, i}$ against the final white dwarf mass $\MWD$.
+The color of the dots indicates how the final orbital period $P_{\mathrm{orb}, f}$ changes with respect to $\lambda=35$ and $\lambda=5$.
+That is, the color indicates the magnitude of $\Delta P = P_{\mathrm{orb}}^{\lambda = 35} - P_{\mathrm{orb}}^{\lambda = 5}$.
+Right panel: scatter plot of $\log_{10} \Delta \Porb$ against the initial orbital period in log scale $\log_{10} P_{\mathrm{orb}, i}$ at the start of RLOF.
+Different colors indicates different kstar type of the primary at the start of RLOF.}
+  \label{dif-kstar-exp}
+\end{figure}
+
+To visualize the difference in final orbital period $\Porb$, and how they corresponds to different \verb|kstar_1| type at the start of RLOF, we create plot the difference in orbital period $\Delta P = P_{\mathrm{orb}}^{\lambda = 35} - P_{\mathrm{orb}}^{\lambda = 5}$ against the white dwarf mass $\MWD$ in Figure \ref{dif-kstar map}.
+From Figure \ref{dif-kstar map}, we can see that for large \verb|kstar_1| type at the start of RLOF, $\Delta P = P_{\mathrm{orb}}^{\lambda = 35} - P_{\mathrm{orb}}^{\lambda = 5}$ is also larger.
+To explain this, we plot the how final white dwarf mass $\MWD$, difference in final orbital period $\Delta \Porb$, and orbital period before mass transfer $P_{\mathrm{orb}, i}$ related to each other in Figure \ref{dif-kstar-exp}.
+We see that the larger initial orbital period $P_{\mathrm{orb}, i}$, the more significant effect does $\lambda$ have on the final orbital period.
+This is not unexpected since during mass transfer, the separation changes as $\dot{a} / a = \text{const}$.
+Hence $\dot{\Porb} / \Porb = \text{const}$.
+Therefore, the larger the initial orbital period, the larger it changes, and thus the larger effect does $\lambda$ have.
+
+\newpage
 \bibliography{reference.bib}

### 2024-07-30 summaries/paper-draft-plain.tex
add population statistics to high mass MT

@@
 Also, we ran these parameters with four different CE efficiencies — $\alphace = 1$, $\alphace = 0.9$, $\alphace = 0.6$, and $\alphace = 0.3$, the same as recorded in \cite{yamaguchi_hi}.
-In total, for each of the $4$ different CE efficiency $\alphace$, we ran a grid of $80000$ binary star systems with different initial separation $a_i$ and CE parameter $\lambda$.
+In total, for each of the $4$ different CE efficiency $\alphace$, we ran a grid of $800000$ binary star systems with different initial separation $a_i$ and CE parameter $\lambda$.
 
@@
 
-A typical change of \verb|evol_type| through the evolution is $\boxed{\text{\textbf{*-3-7-8-4-*-3-7-8-4-*}}}$ (refer to Section \ref{sec:intro} for correspondence between \verb|evol_type| and evolutionary changes), where WD+MS is reached during the second mass transfer process.
-Another possible change of \verb|evol_type| through the volution is $\boxed{\text{\textbf{*-3-7-8-7-8-4-*}}}$.
+Among the 3200000 binary star systems we simulated, a typical change of \verb|evol_type| through the evolution is $\boxed{\text{\textbf{3-7-8-4-3-7-8-4}}}$ (refer to Section \ref{sec:intro} for correspondence between \verb|evol_type| and evolutionary changes, kstar changes and other evolutionary changes before and after the mass transfer are omitted), where WD+MS is reached during the second mass transfer process.
+Another possible change of \verb|evol_type| through the volution is $\boxed{\text{\textbf{3-7-8-7-8-4}}}$.
 It is worth noting that most of these systems experience two mass transfer phases in COSMIC before reaching WD+MS.
 That is, during the evolution process, two CE processes take place, between which there is a period of time.
-Statistically, we found that \textbf{(***TO-DO***)} fraction of all our $320000$ simulated binaries experience two CE processes before reaching WD+MS.
+Statistically, 3146556 of the 3200000 binaries (98\%) we simulated reach WD+MS in their evolutionary process.
+In these 3146556 binaries, 2660287 of them went through two CE processes before reaching WD+MS (85\%).
 To explain this qualitatively, we notice that when the primary loses too much mass during the first CE, the CE process ends, as the radius of the primary becomes smaller than the Roche lobe.
 After the primary fills its Roche lobe and has a large enough envelope again, the CE process resumes and causes two CE processes in total.
+On the contrary, for the rest of the population that reached WD+MS with only one CE, the mass loss of the primary is not enough for the first CE to stop, and the evolution reaches WD+MS directly.
 
@@
 Also, we ran these parameters with four different CE efficiencies — $\alphace = 1$, $\alphace = 0.9$, $\alphace = 0.6$, and $\alphace = 0.3$, the same as recorded in \cite{yamaguchi_lo}.
-In total, for each of the $4$ different CE efficiency $\alphace$, we ran a grid of $80000$ binary star systems with different initial separation $a_i$ and CE parameter $\lambda$.
+In total, for each of the $4$ different CE efficiency $\alphace$, we ran a grid of $800000$ binary star systems with different initial separation $a_i$ and CE parameter $\lambda$.
 
@@
 
-Typical \textbf{pre-spike} and \textbf{spike} systems both follows the evolution path $\boxed{\text{\textbf{*-3-*-5-7-8-4-*-3-*-4-*}}}$ before reacing WD+MS.
+Typical \textbf{pre-spike} and \textbf{spike} systems both follows the evolution path $\boxed{\text{\textbf{3-5-7-8-4-3-4}}}$ before reacing WD+MS.
 Note that the binaries come into contact (\verb|evol_type = 5|), which follows by a common enevelope phase.

### 2024-07-30 summaries/paper-draft-plain.tex
add explanation for stable MT

@@
 \usepackage{amsmath}
+\usepackage{physics}
 \usepackage[labelfont=bf]{caption}
@@
 \newcommand{\au}{\, \mathrm{au}}
+\newcommand{\Md}{M_{\mathrm{d}}}
+\newcommand{\MT}{M_{\mathrm{T}}}
+\newcommand{\Ma}{M_{\mathrm{a}}}
 
@@
 Accretion efficiency is the fraction $\beta$ of all transfered materials that is accreted by the accretor.
-That is, $\dot{M_a} = -\beta\dot{M_d}$ where $M_d$ is the mass of donor and $M_a$ is the mass of accretor.
+That is, $\dot{\Ma} = -\beta\dot{\Md}$ where $\Md$ is the mass of donor and $\Ma$ is the mass of accretor.
 
@@
 Note that the binaries come into contact (\verb|evol_type = 5|), which follows by a common enevelope phase.
-After inspecting the evolutionary process more closely, we found that the second stable mass transfer tends to widens the orbit significantly.
-\textbf{(***TO-DO***)}
+After inspecting the evolutionary process more closely, we found that the second mass transfer (the stable one) tends to widens the orbit significantly.
+
+To qualitatively explain the widening of the orbits, notice that for stable and non-conservative mass transfer with $\dot \Ma = - \beta \dot \Md$, we have
+\[
+  \frac{\dot J}{J} \propto (1 - \beta) q_d \frac{\dot \Md}{\MT},
+\]
+where $q_d = \Md / \Ma$ is the mass ratio between the donor and the accretor. Also,
+\[
+  \frac{\dot a}{a} = 2 \frac{\dot J}{J} - 2 \frac{\dot \Md}{\Md} - 2 \frac{\dot \Ma}{\Ma} + \frac{\dot \MT}{\MT}.
+\]
+Hence, for stable and non-conservative amss transfer, the widening of orbit can be evaluated with
+\begin{equation}
+  \frac{\dot a}{a} \propto - \qty(\frac{2}{q_d} - 2q_d - \beta) \frac{\dot \Md}{\MT}.
+  \label{stableMT}
+\end{equation}
+Also note that the mass transfer begins when the primary is has \verb|evol_type = 8|, a naked Helium star on the Hertzsprung gap, whose wind mass loss is significant.
+For a fast wind mass loss, we have
+\begin{equation}
+  \frac{\dot a}{a} = - \frac{\dot \Md}{\MT}.
+  \label{wind}
+\end{equation}
+This means that strong wind wtripping mass from the primary will cause the orbit to expand even further.
+To illustrate this, we change the \verb|windflag| in \verb|BSEDict| to 0, which correspond to weak wind.
+Now the $a_i = 2.80 \au$ system has final separation $a_f = 0.05 \au$, and the $a_i = 2.90 \au$ system has final separation $a_f = 0.14 \au$, both smaller than the original $0.09 \au$ and $0.27\au$ that correspond to stronger wind.
+
+For our systems, the mass ratio $q_d > 1$, and the value of $q_d$ increases as the initial separation increases.
+Hence, based on Equation \ref{stableMT}, the stable mass transfer process actually shrinks the orbit, while the wind mass loss compensates it by windening the orbit.
+For \textbf{spike} bianries, although $q_d$ is slightly higher, causing the orbit to shrink more according to Equation \ref{stableMT}, the mass loss rate is significantly higher than \textbf{pre-spike} binaries, leading to a more significant increase in orbital separation.
+Also, \textbf{spike} binaries have larger orbital separation $a$ before mass transfer, which also leads to more significant increase in separation during the mass transfer.
+
+For \textbf{merge} binaries, we notice that the separation of binaries at the start of CE is smaller than \textbf{spike} binaries.
+This lack of orbital energy to eject the CE explains the merger event.
+As initial separation contines to incerase, the binaries possess more energy to eject the CE after coming into contact, explaining the reappearance of WD+MS binaries instead of merger.
+
+This concludes our qualitative analysis on the evolution results of $7\Msun + 1\Msun$ through stable mass transfer.
+Again, as shown in Figure \ref{stable_hi}, for stable mass transfer, the observed minimum separation of wide post-mass transfer WD+MS systems $0.15 \au$ can only be achieved in the \textbf{spike} region.
+That is, in our results, all wide post-mass transfer systems are due to the strong wind mass loss from the primary that widens the orbit significantly.
 

### 2024-08-19 summaries/paper-draft-plain.tex
started metisse section

@@
 \newcommand{\Ma}{M_{\mathrm{a}}}
+\newcommand{\Menv}{M_{\mathrm{env}}}
+\DeclareMathOperator{\sgn}{sign}
 
@@
 
+\section{METISSE} \label{sec:metisse}
+To resolve inconsistencies of results across different population synthesis simulation, METISSE, acronym for METhod of Interpolation for Single Star Evolution is developed.
+As documented in documented in \cite{Agrawal_2020,Agrawal_2023}, it is a code package designed for rapidly computing the evolution of stellar populations by interpolating within a set of pre-computed stellar evolutionary tracks.
+To make simulation of binary star systems using METISSE more convenient and improve the simulation accuracy of binary evolution, we modify METISSE to interpolate and save the envelope binding energy of stars.
+In this way, when we need to calculate envelope binding energy in COSMIC, we no longer need to first calculate common envelope $\lambda$ using the fit formula in \cite{claeys2014theoretical}, and then use Equation \ref{ebind} for binding energy.
+Instead, we can use the interpolated binding energy obtained from MESA tracks in METISSE directly, achieving improved accuracy and efficiency of simulation results.
+
+Notice that binding energy is stored in CGS units in the input tracks of METISSE, so the value of energy is very large.
+Hence, it is the best for us to interpolate the log of binding energy per unit envelope mass, instead of binding energy directly.
+Also, for different tracks and different phase of evolution, the binding energy may either be positive or negative.
+To keep the sign of binding energy accessible after interpolation, we preprocess the binding energy field in input tracks into
+\begin{equation}
+  u = \sgn{\Ebind} \cdot \frac{\log_{10} \Ebind}{\Menv}.
+\end{equation}
+Then, we interpolate $u$, just as other parameters are interpolated in METISSE. Finally, for non-degenerate stars, we recover the binding energy
+\begin{equation}
+  \Ebind = \sgn{\hat u} \cdot 10^{\hat u \Menv},
+\end{equation}
+where $\hat u$ is the interpolated value of $u$, and save it in \verb|star_parameters|.
+
 \newpage

### 2024-08-20 summaries/paper-draft-plain.tex
finish adding METISSE to paper draft

@@
 Also, for different tracks and different phase of evolution, the binding energy may either be positive or negative.
-To keep the sign of binding energy accessible after interpolation, we preprocess the binding energy field in input tracks into
+Therefore, to keep the sign of binding energy accessible after interpolation, we preprocess the binding energy field in input tracks into
 \begin{equation}
-  u = \sgn{\Ebind} \cdot \frac{\log_{10} \Ebind}{\Menv}.
+  u = \sgn{\Ebind} \cdot \frac{\log_{10} \abs{\Ebind}}{\Menv},
 \end{equation}
-Then, we interpolate $u$, just as other parameters are interpolated in METISSE. Finally, for non-degenerate stars, we recover the binding energy
+if the binding energy column is provided in the input tracks.
+Then, we interpolate $u$, just as other parameters are interpolated in METISSE.
+Finally, for non-degenerate stars, we recover the binding energy
 \begin{equation}
@@
 \end{equation}
-where $\hat u$ is the interpolated value of $u$, and save it in \verb|star_parameters|.
+where $\hat u$ is the interpolated value of $u$, and save $\Ebind$ in the \verb|binding_energy| field in \verb|star_parameters|.
+In \verb|comenv_lambda.f90|, a file called by COSMIC when common envelope binding enregy is needed, we check if the binding energy column is interpolated.
+If so, we return the value that corresponds to the interpolated $\Ebind$ directly.
+
+For a small caveat, path to input tracks must be provided in file \verb|metisse.input|.
+Both hydrogen tracks and helium tracks is required, and the path must be a full path returned by the \verb|pwd| command in terminal.
+That is, paths should not start with \verb|~/foo|, but something similar to \verb|/Users/john-doe/foo| instead.
+Also, there should not be spaces anywhere in the path, for eaxmple \verb*|Foo/foo\ bar/Bar| for a directory \verb*|foo bar|.
+This will cause the simulation to behave incorrectly.
+
+\begin{figure}
+  \centering
+  \includegraphics[width = 0.6\linewidth]{fig/be-comparison.png}
+  \caption{Comparison of interpolated binding energy using METISSE with binding energy stored in input track for initial mass $M_0 = 5.26 \Msun$.}
+  \label{be-comparison}
+\end{figure}
+
+After modifying the code as described above, we change the initial condition of the simulation in \verb|main.input| and run the program to test it.
+We set the initial mass to be exactly the same initial mass as one input track.
+Then we compare the interpolated binding energy using our method with the binding energy stored in the input track.
+One example of this is shown in Figure \ref{be-comparison}.
+From the figure, we can see that the interpolated binding energy and the values in input MESA tracks has little deviation, validating our approach.
 
